---
title: "Estimating incubation period of novel coronavirus 2019 in Wuhan, China"
output: md_document
editor_options: 
  chunk_output_type: console
---

```{r opts, include=FALSE}
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(echo=F, message=F, warning=F, eval=T,
                      fig.align='center',fig.pos='ht')
```

## Data import

Data freely available at [`data/nCoV-IDD-traveler-data.csv`](https://github.com/HopkinsIDD/ncov_incubation/blob/master/data/nCoV-IDD-traveler-data.csv)) 

We use explicit dates and times when they are reported in the source documents, however when they are not available, we make the following assumptions:

- For cases without a reported right-bound on symptom onset time (SR), we use the time that the case is first presented to a hospital or, lacking that, the time that the source document was published
- For cases without an EL, we use 2019 December 1, which was the onset date for the first reported nCoV-2019 case; though we will test this assumption later
- For cases without an ER, we use the SR
- For cases without an SL, we use the EL


```{r load, include=FALSE}
library(tidyverse)
library(lubridate)
library(coarseDataTools)
library(gridExtra)
# devtools::install_github("reichlab/activemonitr")
library(activeMonitr)
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

set.seed(1)

## read in coronavirus data
ncov_raw <- read_csv("data/nCoV-IDD-traveler-data.csv") %>% 
            rename(EL_date=EL, ER_date=ER, SL_date=SL, SR_date=SR)

## change dates to restrict exposure to after 1 December 2019
## add other times where missing
ncov_ELSR <- ncov_raw %>% 
             # if EL is missing or before 1 Dec 2019, use 1 Dec 2019
             mutate(EL_date=ifelse(is.na(EL_date),"2019-12-01 00:00:00", EL_date) %>% 
                            ymd_hms() %>% 
                            if_else(. < ymd_hms("2019-12-01 00:00:00"), ymd_hms("2019-12-01 00:00:00"), .),
             # if SR is missing, use PR
                    SR_date=ifelse(ymd_hms(SR_date) %>% is.na, PR, SR_date) %>% 
                            ymd_hms(),
           # SR_fever is only for cases with confirmed fever dates
                    SR_fever=ymd_hms(SR_fever))

ncov <- ncov_ELSR %>% 
    # if ER is missing, use SR; if SL is missing, use EL
        mutate(ER_date=if_else(is.na(ER_date), SR_date, ymd_hms(ER_date)),
               SL_date=if_else(is.na(SL_date), EL_date, ymd_hms(SL_date)),
               SL_fever= if_else(is.na(SL_fever) & !is.na(SR_fever), SL_date, ymd_hms(SL_fever))) %>% 
    # calculate days since 1 Dec 2019
        mutate(EL=difftime(EL_date, ymd_hms("2019-12-01 00:00:00"), units="days") %>% 
                  as.numeric(),
               ER=difftime(ER_date, ymd_hms("2019-12-01 00:00:00"), units="days") %>% 
                   as.numeric(),
               SL=difftime(SL_date, ymd_hms("2019-12-01 00:00:00"), units="days") %>% 
                   as.numeric(),
               SR=difftime(SR_date, ymd_hms("2019-12-01 00:00:00"), units="days") %>% 
                   as.numeric(),
               SL_fever=difftime(SL_fever, ymd_hms("2019-12-01 00:00:00"), units="days") %>% 
                   as.numeric(),
               SR_fever=difftime(SR_fever, ymd_hms("2019-12-01 00:00:00"), units="days") %>% 
                   as.numeric(),
              PL=difftime(PL, ymd_hms("2019-12-01 00:00:00"), units="days") %>% 
                  as.numeric(),
              PR=difftime(PR, ymd_hms("2019-12-01 00:00:00"), units="days") %>% 
                  as.numeric(),
              E_int=ER-EL,
              S_int=SR-SL,
              S_fever_int=SR_fever-SL_fever) %>% 
    # remove any entries missing EL, ER, SL, or SR
        filter(!is.na(EL_date), !is.na(ER_date), !is.na(SL_date), !is.na(SR_date)) %>% 
    # remove entries that haven't been reviewed by two people
        filter(!is.na(REVIEWER2)) %>% 
    # remove entries with exposure/onset intervals less than 0
    # remove entries where ER greater than SR or EL greater than SL
        filter(E_int > 0, S_int > 0, ER<=SR, SL>=EL)

## Now lets divide data sets by observation type
## only fevers
ncov_fever <- ncov %>% filter(!is.na(SL_fever) | !is.na(SR_fever))

## only travel outside of China
ncov_foreign <- ncov %>% filter(COUNTRY.DEST != "China" | PROVINCE.DEST %in% c("HongKong", "Macau"))

## only fevers outside of China
ncov_foreign_fever <- ncov_foreign %>% filter(!is.na(SL_fever) | !is.na(SR_fever))

## only cases within mainland China
ncov_mainland <- ncov %>% filter(COUNTRY.DEST == "China" & !(PROVINCE.DEST %in% c("HongKong", "Macau")))

```

## Data summary


```{r data-summary, echo=FALSE, message=FALSE, warning=FALSE}

## sex distribution
table(ncov$SEX)
table(ncov$SEX) / nrow(ncov)
sum(is.na(ncov$SEX))

## age distribution
ncov$age.mid <- (ncov$AGEL + ncov$AGER)/2
summary(ncov$age.mid)

## mainland China vs non mainland
ncov <- ncov %>% 
        mutate(mlChina=ifelse(COUNTRY.DEST=="China" & !PROVINCE.DEST%in%c("HongKong", "Macau"),1,0))

## Regions within mainland China
ncov %>% filter(mlChina==1) %>% group_by(PROVINCE.DEST) %>% summarise(ncases=n()) %>% arrange(desc(ncases))

## Countries + regions outside of mainland China
ncov %>% filter(mlChina==0) %>% group_by(COUNTRY.DEST) %>% summarise(ncases=n()) %>% arrange(desc(ncases))
ncov %>% filter(mlChina==0 & COUNTRY.DEST=="China") %>% group_by(PROVINCE.DEST) %>% summarise(ncases=n()) %>% arrange(desc(ncases))

## Wuhan exposure
ncov$WUHAN_EXP <- if_else(ncov$WUHAN_RESIDENT%in%c("Yes", "Hubei") | ncov$WUHAN_VISITED=="Yes", 1, 0) 
table(ncov$WUHAN_EXP)
prop.table(table(ncov$SEX))
sum(is.na(ncov$SEX))

## Median time from symptom onset to hospitalization
ncov %>% mutate(Smid = (SL + SR)/2,
                Pmid = (PL + PR)/2,
                thosp = Pmid - Smid) %>%
    filter(thosp>=0) %>% select(thosp) %>% summary

## Median exposure interval
summary(ncov$E_int)
summary(ncov$S_int)

```

```{r figure-1, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="This figure displays the exposure and symptom onset windows for each case in our dataset, relative to the right-bound of the exposure window (ER). The blue bars indicate the the exposure windows and the red bars indicate the symptom onset windows for each case. Purple areas are where those two bars overlap."}
dat_sum <- ncov %>%
    mutate(ELnew = EL-ER,
           ERnew = ER-ER,
           Emid = (ELnew + ERnew)/2,
           SLnew = SL-ER,
           SRnew = SR-ER,
           Smid = (SLnew + SRnew)/2,
           UID=reorder(UID, SR-EL))

ggplot(dat_sum, aes(y=factor(UID))) + 
    geom_segment(aes(x=ELnew, xend=ERnew, yend=factor(UID)), 
                 color="#0072B2", size=2, alpha=.25) +
    geom_segment(aes(x=SLnew, xend=SRnew, yend=factor(UID)), 
                 size=2, color="#CC0000", alpha=.25) +
    geom_point(aes(x=Emid, y=factor(UID)), size=0.5, color="#0072B2") +
    geom_point(aes(x=Smid, y=factor(UID)), size=0.5, color="#CC0000") +
    geom_segment(aes(x=Emid, xend=Smid, yend=factor(UID)), size=0.33, color="#999999") +
    #ggtitle("Exposure and symptom onset windows") +
    scale_x_continuous("Days from last possible exposure") +
    scale_y_discrete("Case") +
    theme_bw() +
    theme(axis.text.y = element_blank(),
          axis.ticks.y= element_blank(),
          axis.text.x=element_text(color="black"))
```



## Incubation period estimates

We estimate the incubation period using the coarseDataTools package based on the paper by [Reich *et al*, 2009](https://onlinelibrary.wiley.com/doi/pdf/10.1002/sim.3659).


### Log normal incubation period
We assume a log-normal incubation period and using a bootstrap method for calculating confidence intervals.

The first model we fit is to all of the data and output the median, 2.5th, and 97.5th quantiles (and their confidence intervals):


```{r fit-dic, results='hide', cache=T}
set.seed(1)
#make the data to pass in to dic.fit
ncov_inc_dat <- ncov %>% 
    mutate(type=as.numeric(S_int==0) + as.numeric(E_int==0)) %>% 
    select(EL, ER, SL, SR, type) %>% 
    as.data.frame()

#fit it...use bootstrap interstions for now. MCMC better?
ncov_inc_fit_asym <- dic.fit(ncov_inc_dat,dist="L",
                             ptiles = c(0.025, 0.05, 0.25, 0.5, 0.75, 0.95, 0.975))

ncov_inc_fit_boot <- dic.fit(ncov_inc_dat,dist="L", n.boots=1000,
                             ptiles = c(0.025, 0.05, 0.25, 0.5, 0.75, 0.95, 0.975))
```

```{r fit-dic-gamma, results='hide', cache=T}
set.seed(1)

#fit it...use bootstrap interstions for now. MCMC better?
ncov_gam_fit_boot <- dic.fit(ncov_inc_dat,dist="G", n.boots=1000,
                             ptiles = c(0.025, 0.05, 0.25, 0.5, 0.75, 0.95, 0.975))
```

```{r fit-dic-weibull, results='hide', cache=T}
set.seed(1)

#fit it...use bootstrap interstions for now. MCMC better?
ncov_wei_fit_asym <- dic.fit(ncov_inc_dat,dist="W",
                             ptiles = c(0.025, 0.05, 0.25, 0.5, 0.75, 0.95, 0.975))

ncov_wei_fit_boot <- dic.fit(ncov_inc_dat,dist="W", n.boots=1000,
                             ptiles = c(0.025, 0.05, 0.25, 0.5, 0.75, 0.95, 0.975))
```

```{r fit-dic-erlang, results='hide', cache=T}
set.seed(1)

#fit it...use bootstrap interstions for now. MCMC better?
ncov_erl_fit <- dic.fit.mcmc(ncov_inc_dat,dist="E",
                             ptiles = c(0.025, 0.05, 0.25, 0.5, 0.75, 0.95, 0.975))
```

```{r fit-fever-dics, results='hide', cache=T}
set.seed(1)
#make the data to pass in to dic.fit
ncov_fever_dic <- ncov_fever %>% 
    select(EL, ER, SL=SL_fever, SR=SR_fever) %>% 
    mutate(type=0) %>% 
    as.data.frame()

#fit it...use bootstrap interstions for now. MCMC better?
ncov_fever_fit_asym <- dic.fit(ncov_fever_dic,dist="L",
                             ptiles = c(0.025, 0.05, 0.25, 0.5, 0.75, 0.95, 0.975))

ncov_fever_fit_boot <- dic.fit(ncov_fever_dic,dist="L", n.boots=1000,
                             ptiles = c(0.025, 0.05, 0.25, 0.5, 0.75, 0.95, 0.975))
```

```{r fever-dic-plots, eval=F}
## plot the boot fit and table of intervals
plot(ncov_fever_fit_boot, main="fever-only results")

fever_tbl <- ncov_fever_fit_boot@ests[-c(1:2),-4] %>% 
    as.data.frame()
fever_tbl$diff <- ncov_fever_fit_boot@ests[-c(1:2), "est"] - ncov_inc_fit_boot@ests[-c(1:2), "est"]
knitr::kable(fever_tbl)


## plot the asymptotic fit and table of intervals
# plot(ncov_fever_fit_asym, main="asymptotic results (fever only)")
# ncov_fever_fit_asym
```

```{r fit-foreign-dics, results='hide', cache=T}
set.seed(1)
ncov_foreign_dic <- ncov_foreign %>% 
    select(EL, ER, SL, SR) %>% 
    mutate(type=0) %>% 
    as.data.frame()

ncov_foreign_fit_asym <- dic.fit(ncov_foreign_dic,dist="L",
                             ptiles = c(0.025, 0.05, 0.25, 0.5, 0.75, 0.95, 0.975))

ncov_foreign_fit_boot <- dic.fit(ncov_foreign_dic,dist="L", n.boots=1000,
                             ptiles = c(0.025, 0.05, 0.25, 0.5, 0.75, 0.95, 0.975))

```

```{r foreign-dic-plots, eval=F}
## plot the boot fit and table of intervals
plot(ncov_foreign_fit_boot, main="non-mainland results")

foreign_tbl <- ncov_foreign_fit_boot@ests[-c(1:2),-4] %>% 
    as.data.frame()
foreign_tbl$diff <- ncov_foreign_fit_boot@ests[-c(1:2), "est"] - ncov_inc_fit_boot@ests[-c(1:2), "est"]
knitr::kable(foreign_tbl)

## plot the asymptotic fit and table of intervals
# plot(ncov_foreign_fit_asym, main="asymptotic results (foreign only)")
# ncov_foreign_fit_asym
```

```{r fit-mainland-dics, results='hide', cache=T}
set.seed(1)
ncov_mainland_dic <- ncov_mainland %>% 
    select(EL, ER, SL, SR) %>% 
    mutate(type=0) %>% 
    as.data.frame()

ncov_mainland_fit_asym <- dic.fit(ncov_mainland_dic,dist="L",
                             ptiles = c(0.025, 0.05, 0.25, 0.5, 0.75, 0.95, 0.975))

ncov_mainland_fit_boot <- dic.fit(ncov_mainland_dic,dist="L", n.boots=1000,
                             ptiles = c(0.025, 0.05, 0.25, 0.5, 0.75, 0.95, 0.975))

```

```{r mainland-dic-plots, eval=T}
## plot the boot fit and table of intervals
plot(ncov_mainland_fit_boot, main="non-mainland results")

mainland_tbl <- ncov_mainland_fit_boot@ests[-c(1:2),-4] %>% 
    as.data.frame()
mainland_tbl$diff <- ncov_mainland_fit_boot@ests[-c(1:2), "est"] - ncov_inc_fit_boot@ests[-c(1:2), "est"]
knitr::kable(mainland_tbl)

```

```{r fit-year-dics, results='hide', cache=T}
set.seed(1)
## change dates to restrict exposure to after December 1
## add times where missing
ncov_year_dic <- ncov_raw %>% 
    # if EL is missing or before 1 Dec 2018, use 1 Dec 2018
    mutate(EL_date=ifelse(is.na(EL_date),"2018-12-01 00:00:00", EL_date) %>% 
               ymd_hms() %>% 
               if_else(. < ymd_hms("2018-12-01 00:00:00"),
                       ymd_hms("2018-12-01 00:00:00"), .),
           # if SR is missing, use PR
           SR_date=ifelse(ymd_hms(SR_date) %>% is.na,
                          PR, SR_date) %>% 
               ymd_hms(),
           # SR_fever is only for cases with confirmed fever dates
           SR_fever=ymd_hms(SR_fever)) %>% 
    # if ER is missing, use SR; if SL is missing, use EL
    mutate(ER_date=if_else(is.na(ER_date), SR_date, ymd_hms(ER_date)),
           SL_date=if_else(is.na(SL_date), EL_date, ymd_hms(SL_date)),
           SL_fever= if_else(is.na(SL_fever) & !is.na(SR_fever), 
                             SL_date,
                             ymd_hms(SL_fever))) %>% 
    # calculate days since 1 Dec 2018
    mutate(EL=difftime(EL_date, ymd_hms("2018-12-01 00:00:00"), units="days") %>% 
               as.numeric(),
           ER=difftime(ER_date, ymd_hms("2018-12-01 00:00:00"), units="days") %>% 
               as.numeric(),
           SL=difftime(SL_date, ymd_hms("2018-12-01 00:00:00"), units="days") %>% 
               as.numeric(),
           SR=difftime(SR_date, ymd_hms("2018-12-01 00:00:00"), units="days") %>% 
               as.numeric(),
           SL_fever=difftime(SL_fever,
                             ymd_hms("2018-12-01 00:00:00"), units="days") %>% 
               as.numeric(),
           SR_fever=difftime(SR_fever,
                             ymd_hms("2018-12-01 00:00:00"), units="days") %>% 
               as.numeric(),
           E_int=ER-EL,
           S_int=SR-SL,
           S_fever_int=SR_fever-SL_fever) %>% 
    # remove any entries missing EL, ER, SL, or SR
    filter(!is.na(EL_date), !is.na(ER_date), !is.na(SL_date), !is.na(SR_date)) %>% 
    filter(!is.na(REVIEWER2)) %>% 
    # remove entries with exposure/onset intervals less than 0
    # remove entries where ER greater than SR or EL greater than SL
    # remove entries that haven't been reviewed by two people
    filter(E_int > 0, S_int > 0, ER<=SR, SL>=EL) %>% 
    select(EL, ER, SL, SR) %>% 
    mutate(type=0) %>% 
    as.data.frame()

ncov_year_fit_asym <- dic.fit(ncov_year_dic,dist="L",
                             ptiles = c(0.025, 0.05, 0.25, 0.5, 0.75, 0.95, 0.975))

ncov_year_fit_boot <- dic.fit(ncov_year_dic,dist="L", n.boots=1000,
                             ptiles = c(0.025, 0.05, 0.25, 0.5, 0.75, 0.95, 0.975))

```

```{r year-dic-plots, eval=F}
## plot the boot fit and table of intervals
plot(ncov_year_fit_boot, main="EL-2018 results")

year_tbl <- ncov_year_fit_boot@ests[-c(1:2),-4] %>% 
    as.data.frame()
year_tbl$diff <- ncov_year_fit_boot@ests[-c(1:2), "est"] - ncov_inc_fit_boot@ests[-c(1:2), "est"]
knitr::kable(year_tbl)
```


```{r dic-plots}
## plot the boot fit and table of intervals
ci.col <- rgb(230/255,85/255,13/255,1)

plot(ncov_inc_fit_boot, ylab="Proportion symptomatic cases with symptoms",
     xlab="Days after infection", main="", xlim=c(0,20))
points(y=rep(0.025,2), x=c(ncov_inc_fit_boot@ests['p2.5','CIlow'], ncov_inc_fit_boot@ests['p2.5','CIhigh']), type='l', col=ci.col, lwd=2.5)
points(y=rep(0.5,2), x=c(ncov_inc_fit_boot@ests['p50','CIlow'], ncov_inc_fit_boot@ests['p50','CIhigh']), type='l', col=ci.col, lwd=2.5)
points(y=rep(0.975,2), x=c(ncov_inc_fit_boot@ests['p97.5','CIlow'], ncov_inc_fit_boot@ests['p97.5','CIhigh']), type='l', col=ci.col, lwd=2.5)

knitr::kable(ncov_inc_fit_boot@ests[,-4])

```



## Alternate estimates and sensitivity analyses

### Alternate parameterizations

We fit other commonly-used parameterizations of the incubation period as comparisons to the log-normal distribution: gamma, Weibull, and Erlang. 

```{r other-params}
all_est_plot <- ncov_inc_fit_boot@ests[-2, -4] %>% 
    as.data.frame() %>% 
    mutate(qtile=c("mean", 2.5, 5, 25, 50, 75, 95, 97.5),
           type="log-normal") %>% 
    bind_rows(ncov_gam_fit_boot@ests[-2, -4] %>% 
                  as.data.frame() %>% 
                  # rename(est_fever=est, CIlow_fever=CIlow,
                  #        CIhigh_fever=CIhigh) %>% 
                  mutate(qtile=c("mean",2.5, 5, 25, 50, 75, 95, 97.5),
                         type="gamma")) %>% 
    bind_rows(ncov_wei_fit_boot@ests[-2, -4] %>% 
                  as.data.frame() %>% 
                  # rename(est_foreign=est, CIlow_foreign=CIlow,
                  #        CIhigh_foreign=CIhigh) %>% 
                  mutate(qtile=c("mean", 2.5, 5, 25, 50, 75, 95, 97.5),
                         type="weibull")) %>% 
    bind_rows(ncov_erl_fit@ests[-2, -4] %>% 
                  as.data.frame() %>% 
                  # rename(est_year=est, CIlow_year=CIlow,
                  #        CIhigh_year=CIhigh) %>% 
                  mutate(qtile=c("mean", 2.5, 5, 25, 50, 75, 95, 97.5),
                         type="erlang")) %>% 
    mutate(est=ifelse(qtile=="mean", exp(est), est),
           CIlow=ifelse(qtile=="mean", exp(CIlow), CIlow),
           CIhigh=ifelse(qtile=="mean", exp(CIhigh), CIhigh),
           study="JHU-IDD",
           virus="nCoV") %>% 
    # bind_rows(other_studies) %>% 
    filter(qtile %in% c(2.5, 50, 97.5))# %>% 
    # select(-CIhigh, -CIlow) %>% 
    # spread(qtile, est, sep="_")

ggplot(data=all_est_plot,
       aes(y=est, ymin=CIlow, ymax=CIhigh, x=as.factor(qtile), color=type)) +
    geom_errorbar(height=0.2, position=position_dodge(0.9)) +
    geom_point(position=position_dodge(0.9)) +
    scale_y_continuous("Incubation time, in days (with 95% CIs)", limits=c(0,16)) +
    scale_x_discrete("Estimate quantile") +
    scale_color_manual("Est\ntype",
                       values=cbbPalette[c(2,6,1,4)]) +
    theme_bw() + coord_flip() +
    theme(axis.text=element_text(color="black"))

dic_lls <- tibble(dist=c("log-normal",
                         "gamma",
                         "Weibull",
                         "Erlang"),
                  ll=c(ncov_inc_fit_boot@loglik,
                       ncov_gam_fit_boot@loglik,
                       ncov_wei_fit_boot@loglik,
                       ncov_erl_fit@loglik))

ncov_gam_fit_boot
ncov_wei_fit_boot
ncov_erl_fit
```

The median estimates are very similar across parameterizations, while the Weibull distribution has a slightly smaller value at the 2.5th percentile and the log-normal distribution has a slightly larger value at the 97.5th percentile.
The log-likelihoods were very similar between distributions; the `r dic_lls$dist[which.max(dic_lls$ll)]` distribution having the largest log-likelihood (`r max(dic_lls$ll) %>% round(2)`) and the `r dic_lls$dist[which.min(dic_lls$ll)]` distribution having the smallest log-likelihood (`r min(dic_lls$ll) %>% round(2)`).

The gamma distribution has an estimated shape parameter of `r ncov_gam_fit_boot@ests["shape","est"] %>% round(2)` (95% CI: `r ncov_gam_fit_boot@ests["shape","CIlow"] %>% round(2)`-`r ncov_gam_fit_boot@ests["shape","CIhigh"] %>% round(2)`) and a scale parameter of `r ncov_gam_fit_boot@ests["scale","est"] %>% round(2)` (95% CI: `r ncov_gam_fit_boot@ests["scale","CIlow"] %>% round(2)`-`r ncov_gam_fit_boot@ests["scale","CIhigh"] %>% round(2)`).
The Weibull distribution has an estimated shape parameter of `r ncov_wei_fit_boot@ests["shape","est"] %>% round(2)` (95% CI: `r ncov_wei_fit_boot@ests["shape","CIlow"] %>% round(2)`-`r ncov_wei_fit_boot@ests["shape","CIhigh"] %>% round(2)`) and a scale parameter of `r ncov_wei_fit_boot@ests["scale","est"] %>% round(2)` (95% CI: `r ncov_wei_fit_boot@ests["scale","CIlow"] %>% round(2)`-`r ncov_wei_fit_boot@ests["scale","CIhigh"] %>% round(2)`).
The Erlang distribution has an estimated shape parameter of `r ncov_erl_fit@ests["shape","est"] %>% round(2)` (95% CI: `r ncov_erl_fit@ests["shape","CIlow"] %>% round(2)`-`r ncov_erl_fit@ests["shape","CIhigh"] %>% round(2)`) and a scale parameter of `r ncov_erl_fit@ests["scale","est"] %>% round(2)` (95% CI: `r ncov_erl_fit@ests["scale","CIlow"] %>% round(2)`-`r ncov_erl_fit@ests["scale","CIhigh"] %>% round(2)`).

### Sensitivity analyses

To make sure that our overall incubation estimates are sound, we ran a few analyses on subsets to see if the results held up.
Since the winter often brings cold air and other pathogens that can cause sore throats and coughs, we ran an analysis using only cases that reported a fever.
Since a plurality of our cases came from mainland China, where assumptions about local transmission may be less firm, we ran an analysis without those cases.
Finally, we challenge our assumption that unknown ELs can be assumed to be 2019 December 1 ([Nextstrain estimates that it could have happened as early as September](https://nextstrain.org/ncov?dmax=2019-12-04&m=num_date)), by setting unknown ELs to 2018 December 1.

```{r all-sens-plot, warning=F, message=F}
all_sens_plot <- ncov_inc_fit_boot@ests[-2, -4] %>% 
    as.data.frame() %>% 
    mutate(qtile=c("mean", 2.5, 5, 25, 50, 75, 95, 97.5),
           type="all") %>% 
    bind_rows(ncov_fever_fit_boot@ests[-2, -4] %>% 
                  as.data.frame() %>% 
                  # rename(est_fever=est, CIlow_fever=CIlow,
                  #        CIhigh_fever=CIhigh) %>% 
                  mutate(qtile=c("mean",2.5, 5, 25, 50, 75, 95, 97.5),
                         type="fever")) %>% 
    bind_rows(ncov_foreign_fit_boot@ests[-2, -4] %>% 
                  as.data.frame() %>% 
                  # rename(est_foreign=est, CIlow_foreign=CIlow,
                  #        CIhigh_foreign=CIhigh) %>% 
                  mutate(qtile=c("mean", 2.5, 5, 25, 50, 75, 95, 97.5),
                         type="non-mainland")) %>% 
    bind_rows(ncov_mainland_fit_boot@ests[-2, -4] %>% 
                  as.data.frame() %>% 
                  # rename(est_year=est, CIlow_year=CIlow,
                  #        CIhigh_year=CIhigh) %>% 
                  mutate(qtile=c("mean", 2.5, 5, 25, 50, 75, 95, 97.5),
                         type="mainland")) %>% 
    mutate(est=ifelse(qtile=="mean", exp(est), est),
           CIlow=ifelse(qtile=="mean", exp(CIlow), CIlow),
           CIhigh=ifelse(qtile=="mean", exp(CIhigh), CIhigh),
           study="JHU-IDD",
           virus="nCoV") %>% 
    # bind_rows(other_studies) %>% 
    filter(qtile %in% c(2.5, 50, 97.5))# %>% 
    # select(-CIhigh, -CIlow) %>% 
    # spread(qtile, est, sep="_")

all_sens_plot <- ggplot(data=all_sens_plot,
                        aes(y=est, ymin=CIlow, ymax=CIhigh, x=as.factor(qtile), color=type)) +
    geom_errorbar(height=0.2, position=position_dodge(0.9)) +
    geom_point(position=position_dodge(0.9)) +
    scale_y_continuous("Incubation time, in days (with 95% CIs)", limits=c(0,21)) +
    scale_x_discrete("Estimate quantile") +
    scale_color_manual("Est\ntype",
                       values=cbbPalette[c(1,6,4,7)]) +
    ggtitle("A") +
    theme_bw() + coord_flip() +
    theme(axis.text=element_text(color="black"))
```

## Comparison to Backer

[Backer, Klinkenberg, & Wallinga](https://www.medrxiv.org/content/10.1101/2020.01.27.20018986v1.full.pdf+html) estimated the incubation windows based on 34 early nCoV cases that traveled from Wuhan to other regions in China.
A comparison of our incubation windows are shown below:

```{r comparison}
backer_params <- read_csv("data/backer-params.csv")
backer_comp <- ncov_inc_fit_boot@ests[-2, -4] %>% 
    as.data.frame() %>% 
    mutate(qtile=c("mean",2.5, 5, 25, 50, 75, 95, 97.5),
           type="log-normal") %>% 
    bind_rows(ncov_gam_fit_boot@ests[-2, -4] %>%
                  as.data.frame() %>%
                  mutate(qtile=c("mean",2.5, 5, 25, 50, 75, 95, 97.5),
                         type="gamma")) %>%
    bind_rows(ncov_wei_fit_boot@ests[-2, -4] %>% 
                  as.data.frame() %>% 
                  mutate(qtile=c("mean",2.5, 5, 25, 50, 75, 95, 97.5),
                         type="weibull")) %>% 
    bind_rows(ncov_erl_fit@ests[-2, -4] %>%
                  as.data.frame() %>%
                  mutate(qtile=c("mean",2.5, 5, 25, 50, 75, 95, 97.5),
                         type="erlang")) %>%
    mutate(est=ifelse(qtile=="mean", exp(est), est),
           CIlow=ifelse(qtile=="mean",exp(CIlow), CIlow),
           CIhigh=ifelse(qtile=="mean",exp(CIhigh), CIhigh),
           study="JHU-IDD",
           virus="nCoV") %>% 
    filter(qtile %in% c(2.5, 25, 50, 75, 97.5)) %>% 
    select(-CIhigh, -CIlow) %>%
    spread(qtile, est, sep="_") %>%
    bind_rows(backer_params %>% 
                  mutate(`qtile_2.5`=ifelse(type=="weibull",
                                          qweibull(.025, par1, par2),
                                          ifelse(type=="gamma",
                                                 qgamma(.025, par1, scale=par2),
                                                 qlnorm(.025, par1, par2))),
                         qtile_25=ifelse(type=="weibull",
                                          qweibull(.25, par1, par2),
                                          ifelse(type=="gamma",
                                                 qgamma(.25, par1, scale=par2),
                                                 qlnorm(.25, par1, par2))),
                         qtile_50=ifelse(type=="weibull",
                                          qweibull(.5, par1, par2),
                                          ifelse(type=="gamma",
                                                 qgamma(.5, par1, scale=par2),
                                                 qlnorm(.5, par1, par2))),
                         qtile_75=ifelse(type=="weibull",
                                          qweibull(.75, par1, par2),
                                          ifelse(type=="gamma",
                                                 qgamma(.75, par1, scale=par2),
                                                 qlnorm(.75, par1, par2))),
                         `qtile_97.5`=ifelse(type=="weibull",
                                          qweibull(.975, par1, par2),
                                          ifelse(type=="gamma",
                                                 qgamma(.975, par1, scale=par2),
                                                 qlnorm(.975, par1, par2)))) %>% 
                  select(-par1, -par2)) %>% 
    mutate(study_order=ifelse(study=="JHU-IDD", 3,
                              ifelse(study=="Backer 2020", 2, 1)),
           study=reorder(study, study_order))

comparison_plot <- ggplot(data=backer_comp,
                          aes(middle=qtile_50, ymin=qtile_2.5, ymax=qtile_97.5,
                              lower=qtile_25, upper=qtile_75, x=type,
                              fill=study)) +
    geom_boxplot(stat="identity",
                 position = position_dodge2(preserve = "single")) +
    scale_y_continuous("Incubation time, in days",
                       limits=c(0,16)) +
    scale_x_discrete("Distribution") +
    scale_fill_manual("Study",
                      breaks=c("JHU-IDD", "Backer 2020", "Li 2020"),
                      values=cbbPalette[c(4,2,3)]) +
    ggtitle("B") +
    theme_bw() + coord_flip() +
    theme(axis.text=element_text(color="black"))

grid.arrange(all_sens_plot, comparison_plot)
```

```{r missed-cases}
## Figure 1
## plot with x-axis in days, y axis as cases missed per XX monitored 
## (similar to fig 3 of activemonitr manuscript)

ncov_gamma_dat <- tibble(shape=ncov_gam_fit_boot@samples$par1,
                         scale=ncov_gam_fit_boot@samples$par2,
                         idx=seq(nrow(ncov_gam_fit_boot@samples))) %>% 
    mutate(median=qgamma(0.5, shape=shape, scale=scale),
           p95=qgamma(0.95, shape=shape, scale=scale),
           chain=1)

phis <- c(1/100, 1/1000, 1/10000)
durs <- 1:25
yrange <- c(1e-18, 1e-2)

## determine max u for each disease such that max(u)/2 + m = 90th percentile of T
tmp_ncov <- mean(qgamma(.90, shape=ncov_gamma_dat$shape, 
                        scale=ncov_gamma_dat$scale))
maxu_ncov <- 2*(tmp_ncov - mean(ncov_gamma_dat$median))

## make plots/data
ncov_monitor_probs <- plot_risk_uncertainty(ncov_gamma_dat,# phi=phis, 
                                            max_u = maxu_ncov,
                                            durations=durs, 
                                            return_plot=FALSE,
                                            return_data=TRUE,
                                            include_xlab = FALSE, yrange=yrange, 
                                            include_legend=FALSE)

ncov_monitor_cases <- ncov_monitor_probs$data %>% 
    mutate(
        escaped_cases_per_1k_p05 = p05 * 1000,
        escaped_cases_per_1k_p25 = p25 * 1000,
        escaped_cases_per_1k_p50 = p50 * 1000,
        escaped_cases_per_1k_p75 = p75 * 1000,
        escaped_cases_per_1k_p95 = p95 * 1000,
        risk_group = factor(phi, levels=phis,
                            labels=c("high risk (1 in 100)",
                                     "some risk (1 in 1,000)",
                                     "low risk (1 in 10,000)"))
    )

ggplot(ncov_monitor_cases,
       aes(x=d, y=escaped_cases_per_1k_p50, color=risk_group,
           group=risk_group, fill=risk_group)) +
    # geom_ribbon(aes(ymin=escaped_cases_per_1k_p25,
                    # ymax=escaped_cases_per_1k_p75), alpha=.5) +
    geom_ribbon(aes(ymin=escaped_cases_per_1k_p05,
                    ymax=escaped_cases_per_1k_p95), alpha=.25) +
    geom_line(size=2) +
    ylab("expected cases missed per 1,000 monitored") +
    xlab("duration of monitoring (days)") +
    scale_color_brewer(type="qual") + 
    scale_fill_brewer(type="qual") + 
    theme_bw() +
    # scale_y_log10()+
    theme(legend.position = c(.8,.8), legend.justification = c(1,1))
```

