---
title: "The estimation of the Wuhan coronavirus incubation time"
output: md_document
---

```{r opts, include=FALSE}
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(echo=F, message=T, warning=T, eval=T,
                      fig.align='center',fig.pos='ht')
```

Analysis based on similar procedures for determining the incubation period of Zika virus (Lessler *et al*, 2016.

```{r load, include=FALSE}
library(tidyverse)
library(lubridate)
library(coarseDataTools)
library(gridExtra)
library(rstan)

set.seed(1)

## read in coronavirus data
ncov_raw <- read_csv("data/nCoV-IDD-traveler-data.csv") %>% 
    rename(EL_date=EL, ER_date=ER, SL_date=SL, SR_date=SR)

## change dates to restrict exposure to after December 1
## add times where missing
ncov_ELSR <- ncov_raw %>% 
    # if EL is missing or before 1 Dec 2019, use 1 Dec 2019
    mutate(EL_date=ifelse(is.na(EL_date),"2019-12-01 00:00:00", EL_date) %>% 
               ymd_hms() %>% 
               if_else(. < ymd_hms("2019-12-01 00:00:00"),
                       ymd_hms("2019-12-01 00:00:00"), .),
           # if SR is missing, use PR
           SR_date=ifelse(ymd_hms(SR_date) %>% is.na,
                          PR, SR_date) %>% 
               ymd_hms(),
           # SR_fever is only for cases with confirmed fever dates
           SR_fever=ymd_hms(SR_fever))

ncov <- ncov_ELSR %>% 
    # if ER is missing, use SR; if SL is missing, use EL
    mutate(ER_date=if_else(is.na(ER_date), SR_date, ymd_hms(ER_date)),
           SL_date=if_else(is.na(SL_date), EL_date, ymd_hms(SL_date)),
           SL_fever= if_else(is.na(SL_fever) & !is.na(SR_fever), 
                             SL_date,
                             ymd_hms(SL_fever))) %>% 
    # calculate days since 1 Dec 2019
    mutate(EL=difftime(EL_date, ymd_hms("2019-12-01 00:00:00"), units="days") %>% 
               as.numeric(),
           ER=difftime(ER_date, ymd_hms("2019-12-01 00:00:00"), units="days") %>% 
               as.numeric(),
           SL=difftime(SL_date, ymd_hms("2019-12-01 00:00:00"), units="days") %>% 
               as.numeric(),
           SR=difftime(SR_date, ymd_hms("2019-12-01 00:00:00"), units="days") %>% 
               as.numeric(),
           SL_fever=difftime(SL_fever,
                             ymd_hms("2019-12-01 00:00:00"), units="days") %>% 
               as.numeric(),
           SR_fever=difftime(SR_fever,
                             ymd_hms("2019-12-01 00:00:00"), units="days") %>% 
               as.numeric(),
           E_int=ER-EL,
           S_int=SR-SL,
           S_fever_int=SR_fever-SL_fever) %>% 
    # remove any entries missing EL, ER, SL, or SR
    filter(!is.na(EL_date), !is.na(ER_date), !is.na(SL_date), !is.na(SR_date)) %>% 
    filter(!is.na(REVIEWER2)) %>% 
    # remove entries with exposure/onset intervals less than 0
    # remove entries where ER greater than SR or EL greater than SL
    # remove entries that haven't been reviewed by two people
    filter(E_int > 0, S_int > 0, ER<=SR, SL>=EL)# %>% 
    # mutate(SL=ifelse(SL==EL, SL+.001, SL),
    #        ER=ifelse(ER==SR, ER-.001, ER))

## Now lets divide data sets by observation type
## only fevers
ncov_fever <- ncov %>% 
    filter(!is.na(SL_fever) | !is.na(SR_fever))

## only travel outside of China
ncov_foreign <- ncov %>% 
    filter(COUNTRY.DEST != "China" | PROVINCE.DEST %in% c("HongKong", "Macau"))

## only fevers outside of China
ncov_foreign_fever <- ncov_foreign %>% 
    filter(!is.na(SL_fever) | !is.na(SR_fever))
```

```{r data-summary, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7}
dat_sum <- ncov %>%
    mutate(ELnew = EL-ER,
           ERnew = ER-ER,
           SLnew = SL-ER,
           SRnew = SR-ER)

ggplot(dat_sum, aes(y=factor(UID))) + 
    geom_segment(aes(x=ELnew, xend=ERnew, yend=factor(UID)), 
                 color="blue", size=2) +
    geom_segment(aes(x=SLnew, xend=SRnew, yend=factor(UID)), 
                 size=2, color="red", alpha=.5) +
    #ggtitle("Incubation Period data") + 
    scale_x_continuous("") +
    ylab("UID") +
    theme_bw() +
    # coord_cartesian(xlim = c(-40, 50)) +
    theme(axis.text.y = element_text(size=6))

## number of fevers
num_fevers <- nrow(ncov_fever)
## number of foreign cases
num_foreign <- nrow(ncov_foreign)
## number of males females and unknown
table(ncov$SEX %>% as.factor)
## age ranges
quantile((ncov$AGEL+ncov$AGER)/2, probs=c(.25, .5, .75), na.rm=T)

```

We have data with which to estimate the incubation period from `r nrow(ncov)` people.
The median exposure interval was `r round(median(ncov$E_int),1)` (range: `r round(min(ncov$E_int),1)`-`r round(max(ncov$E_int),1)`).
The median symptom onset interval was `r round(median(ncov$S_int),1)` (range: `r round(min(ncov$S_int),1)`-`r round(max(ncov$S_int),1)`).

# coarseDataTools Bootstrap based analyses

We first estimate each relative distribution using coarseDataTools and bootstrapping. Replicating main analysis, assuming a log-normal distribution for the incubation period, and Weibull for the other relevant distributions. 

## Incubation Period

Basic incubation period analysis assuming a log-normal incubation period and using a bootstrap method for calculating confidence intervals.
We also run the asymptotic intervals, which run much faster.

```{r fit-dic, results='hide', cache=T}
#make the data to pass in to dic.fit
ncov_inc_dat <- ncov %>% 
    mutate(type=as.numeric(S_int==0) + as.numeric(E_int==0)) %>% 
    select(EL, ER, SL, SR, type) %>% 
    as.data.frame()

#fit it...use bootstrap interstions for now. MCMC better?
ncov_inc_fit_asym <- dic.fit(ncov_inc_dat,dist="L",
                             ptiles = c(0.025, 0.05, 0.25, 0.5, 0.75, 0.95, 0.975))

ncov_inc_fit_boot <- dic.fit(ncov_inc_dat,dist="L", n.boots=1000,
                             ptiles = c(0.025, 0.05, 0.25, 0.5, 0.75, 0.95, 0.975))
```

```{r dic-plots}
## plot the boot fit and table of intervals
plot(ncov_inc_fit_boot, main="bootstrap results")
ncov_inc_fit_boot

## plot the asymptotic fit and table of intervals
plot(ncov_inc_fit_asym, main="asymptotic results")
ncov_inc_fit_asym

```

We also made incubation estimates for the fever-only and foreign-only subsets.

```{r fit-fever-dics, results='hide', cache=T}
#make the data to pass in to dic.fit
ncov_fever_dic <- ncov_fever %>% 
    select(EL, ER, SL=SL_fever, SR=SR_fever) %>% 
    mutate(type=0) %>% 
    as.data.frame()

#fit it...use bootstrap interstions for now. MCMC better?
ncov_fever_fit_asym <- dic.fit(ncov_fever_dic,dist="L",
                             ptiles = c(0.025, 0.05, 0.25, 0.5, 0.75, 0.95, 0.975))

ncov_fever_fit_boot <- dic.fit(ncov_fever_dic,dist="L", n.boots=1000,
                             ptiles = c(0.025, 0.05, 0.25, 0.5, 0.75, 0.95, 0.975))
```

```{r fever-dic-plots}
## plot the boot fit and table of intervals
plot(ncov_fever_fit_boot, main="bootstrap results (fever only)")
ncov_fever_fit_boot

## plot the asymptotic fit and table of intervals
plot(ncov_fever_fit_asym, main="asymptotic results (fever only)")
ncov_fever_fit_asym
```

```{r fit-foreign-dics, results='hide', cache=T}
ncov_foreign_dic <- ncov_foreign %>% 
    select(EL, ER, SL, SR) %>% 
    mutate(type=0) %>% 
    as.data.frame()

ncov_foreign_fit_asym <- dic.fit(ncov_foreign_dic,dist="L",
                             ptiles = c(0.025, 0.05, 0.25, 0.5, 0.75, 0.95, 0.975))

ncov_foreign_fit_boot <- dic.fit(ncov_foreign_dic,dist="L", n.boots=1000,
                             ptiles = c(0.025, 0.05, 0.25, 0.5, 0.75, 0.95, 0.975))

```

```{r foreign-dic-plots}
## plot the boot fit and table of intervals
plot(ncov_foreign_fit_boot, main="bootstrap results (foreign only)")
ncov_foreign_fit_boot

## plot the asymptotic fit and table of intervals
plot(ncov_foreign_fit_asym, main="asymptotic results (foreign only)")
ncov_foreign_fit_asym
```

```{r fit-foreign-fever-dics, results='hide', cache=T, eval=F}
ncov_foreign_fever_dic <- ncov_foreign_fever %>% 
    select(EL, ER, SL, SR) %>% 
    mutate(type=0) %>% 
    as.data.frame()

## foreign fever not working
ncov_foreign_fever_fit_asym <- dic.fit(ncov_foreign_fever_dic,dist="L",
                             ptiles = c(0.025, 0.05, 0.25, 0.5, 0.75, 0.95, 0.975))
ncov_foreign_fever_fit_boot <- dic.fit(ncov_foreign_fever_dic,dist="L",
                                       n.boots=1000,
                             ptiles = c(0.025, 0.05, 0.25, 0.5, 0.75, 0.95, 0.975))

```

```{foreign-fever-dics, eval=F}
## plot the asymptotic fit and table of intervals
plot(ncov_foreign_fever_fit_asym, main="asymptotic results (foreign only)")
ncov_foreign_fever_fit_asym
```

```{r stan-incubation, eval=FALSE}
ncov_model <- stan_model("source/DistributionFitL.stan")

inc_est <- sampling(ncov_model,
                    data=list(N=nrow(ncov),
                              EL=ncov$EL,
                              ER=ncov$ER,
                              SL=ncov$SL,
                              SR=ncov$SR),
                    chains=1,
                    iter=1000,
                    warmup=500,
                    control=list(adapt_delta=0.8,
                                 max_treedepth=10),
                    seed=1)
```
