---
title: "Supplement to Time to Key Events in the Course of Zika Infection and their Implications for Surveillance"
author: "Justin Lessler, Jacob M. Konikoff, Cassandra T. Ott, Andrea C. Carcelen, Joe Williamson, Qifang Bi, Nicholas G. Reich, Derek A. T. Cummings, Lauren M. Kucirka, Lelia H. Chaisson"
date: "Monday, February 29, 2016"
output: word_document
references:
- id: brooks1998
  title: General Methods for Monitoring Convergence of Iterative Simulations
  author:
  - family: Brooks
    given: S.P.
  - family: Gelman
    given: A.G.
  container-title: Journal of Computational and Graphical Statistics
  volume: 7
  DOI: 10.1080/10618600.1998.10474787
  issue: 4
  page: 435-455
  type: article-journal
  issued:
    year: 1998
- id: reich2009
  title: Estimating incubation period distributions with coarse data
  author:
    - family: Reich
      given: N.G.
    - family: Lessler
      given: J.
    - family: Cummings
      given: D.A.T.
    - family: Brookmeyer
      given: R.
  container-title: Statisics in Medicine
  volume: 28
  page: 2769-2784
  issued:
    year: 2009
- id: lessler2009
  title: Incubation periods of acute respiratory viral infections a systematic review.
  author:
    - family: Lessler
      given: J.
    - family: Reich
      given: N.G.
    - family: Brookmeyer
      given: R.
    - family: Perl
      given: T.M.
    - family: Nelson
      given: K.E.
    - family: Cummings
      given: D.A.T.
  container-title: Lancet Infectious Disease
  volume: 9
  page: 291-300
  issued:
    year: 2009
---


```{r, echo=FALSE, warning=FALSE,message=FALSE}
#libraries 
library(coda)
````

```{r, echo=FALSE}
#Variables to make code work with the local environemnt


````

# 1 Statistical Methods

## 1.1 Data cleaning 

All analytic data sets and code, including the R-markdown file used to generate this supplement, is available from https://github.com/HopkinsIDD/ZikaLitReview. 

Analytic data is available in the file ZikaLitReviewData.csv, which consists of one line per observation of a bounded time of symptom onset, virologic testing or serologic testing. The main analysis is based on all abstracted observations, minus those that were reports of perinatal transmission. 


```{r, echo=FALSE}
zika <- read.csv("ZikaLitReviewData.csv")

#DROP PERINATAL TRANSMISSIONS
zika <- zika[which(zika$TRANS.MECH!="perinatal"),]  

#Now lets divide data sets by observation type.

#Incubation period
zika.inc <- zika[!is.na(zika$SL),]

#Period of viral shedding
zika.vshed <- zika[!is.na(zika$VL),]

#Time to seropositiveity
zika.sero <- zika[!is.na(zika$AL),]
```

### 1.1.1. Determining possible times of viral clearance
To calculate the period of viral clearance from blood, defined as the last moment there is detectable Zika virus in a blood sample, we use the following procedure:

1. All virolgic tests from samples other than blood are excluded.
2. The earliest possible time of viral clearance is:
    + *if a positive or equivocal test was reported:* the time of the last postive virologic test
    + *if no positive or equivoval test was reported:* the earliest possible time of exposure
3. The latest possible time of viral clearance is:
    + *if negative and positive tests were reported:* the time of the first negative test following a positive test
    + *if only negatve tests were reported:* the time of the earliest negative test
    + *if no negative tests were reported:* infinite/undefined



```{r, echo=FALSE, warning=FALSE}
#first pass, restrict to blood (i.e., not saliva or urine)
zika.vshed.bl <- zika.vshed[which(zika.vshed$SAMPLE.TYPE %in%c("serum","blood")),]

#Go through and find the start of the latest period in which
#someone showed viral shedding and the start of the subsequent
#period where they did not....these bound the possible time of viral clearance
for (uid in unique(zika.vshed.bl$UID)) {
  inds <- which(zika.vshed.bl$UID==uid)
  tmp <- zika.vshed.bl[inds,] 
  
  #taking greater than or equal to 1 means we will treat equivocal results as 
  #positive
  low.time <- max(tmp$VL[tmp$SHEDDING>=1])
  
  #no earliest point...go to earliest time of infection
  if (low.time<tmp$EL[1]) {low.time <- tmp$EL[1]}
  
  high.time <- min(tmp$VR[tmp$SHEDDING==0 &  tmp$VR>low.time])

  zika.vshed.bl$VCL[inds] <- low.time
  zika.vshed.bl$VCR[inds] <- high.time
  }

  #make the analytic dataset
  vshed.bl.anal <- zika.vshed.bl[,c("UID","EL","ER","VCL","VCR")]
  vshed.bl.anal$type <- 0
  vshed.bl.anal <- unique(vshed.bl.anal) #one row per observation
  
```


### 1.1.2. Determining possible times of seroconversion

Seroconversion was defined as the time at which antibody first becomes detectable in the blood (level of detection defined by the original article). The period of possible seroconversion is calculated similarly to the period of possible viral clearance:

1. The earliest possible time of seroconversion is:
    + *if a negative test was reported:* the time of the last negative serologic test
    + *if no negative test was reported:* the earliest possible time of exposure
3. The latest possible time of seroconversion is:    
    + *if a postive or equivocal serological test is reported:* the time of the first positive or equivoval test
    + *if no postive or equivcval tests were reported:* infinite/undefined


```{r, echo=FALSE, warning=FALSE}
# For the first pass we will assume any seropositivity indicates
#serpositivity by IgG

zika.sero$ACL <- NA
zika.sero$ACR <- NA

#for the purposes of this analysis assume that a
#NA in any antibody should be replaces with the IgM value
zika.sero$ANYANTIBODY[is.na(zika.sero$ANYANTIBODY)] <- 
  zika.sero$IGM[is.na(zika.sero$ANYANTIBODY)]


#Loop through unique IDs bounding their time of 
#seroconversion
for (uid in unique(zika.sero$UID)) {
  inds <- which(zika.sero$UID == uid)
  tmp <- zika.sero[inds,]
  
  
  # print(inds) # DEBUG
  
  #lowest time is the latest time with a negative serology
  low.time <- max(tmp$AL[tmp$ANYANTIBODY==0])
  
   
  #no earliest point...go to earliest possible  time of infection
  if (low.time<tmp$EL[1]) {low.time <- tmp$EL[1]}
  
  #latest possible time is the earliest time of a positive antibody test. For our
  #purposes, assume that equivocal is seropositive
  high.time <- min(tmp$AR[which(tmp$ANYANTIBODY>0)])
  
  
  zika.sero$ACL[inds] <- low.time
  zika.sero$ACR[inds] <- high.time
  
}

#make an analytic dataset with one row per person
sero.anal <- zika.sero[,c("UID", "EL", "ER", "ACL","ACR")]
sero.anal$type <- 0 
sero.anal <- unique(sero.anal)
```


```{r, echo=FALSE}
#Merges evertying into a single dataset


#Create a dataset with everything lined up for passing into STAN
zika.anal <- zika.inc[,c("UID","EL","ER","SL","SR")]
zika.anal <- merge(zika.anal, vshed.bl.anal, all=TRUE)
zika.anal$type <- 0 #we don't use this, but helps with the merge.
zika.anal <- merge(zika.anal, sero.anal, all=TRUE)

#renormalize everything so EL is 0
zika.anal[,c("EL","ER","SL","SR","VCL","VCR","ACL","ACR")] <-
  zika.anal[,c("EL","ER","SL","SR","VCL","VCR","ACL","ACR")] -zika.anal$EL  

 

```

### 1.1.3 Visual summary of interval-censored data
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7}
library(dplyr)
library(ggplot2)
library(gridExtra)
dat <- zika.anal %>%
    mutate(ELnew = EL-ER,
           ERnew = ER-ER,
           SLnew = SL-ER,
           SRnew = SR-ER,
           VCLnew = VCL-ER,
           VCRnew = VCR-ER,
           ACLnew = ACL-ER,
           ACRnew = ACR-ER)

IPplot <- ggplot(dat, aes(y=factor(UID))) + 
    geom_segment(aes(x=ELnew, xend=ERnew, yend=factor(UID)), 
                 color="blue", size=2) +
    geom_segment(aes(x=SLnew, xend=SRnew, yend=factor(UID)), 
                 size=2, color="red", alpha=.5) +
    #ggtitle("Incubation Period data") + 
    xlab(NULL) +
    ylab("UID") +
    coord_cartesian(xlim = c(-40, 50)) +
    theme(axis.text.y = element_text(size=6)) +
    annotate("text", x=-35, y="900.1", label="A")

VCplot <- ggplot(dat, aes(y=factor(UID))) + 
    geom_segment(aes(x=ELnew, xend=ERnew, yend=factor(UID)), 
                 color="blue", size=2) +
    geom_segment(aes(x=VCLnew, xend=VCRnew, yend=factor(UID)), 
                 color="red", alpha=.5, size=2)  +
    #ggtitle("Viral shedding data") + 
    xlab(NULL) +
    ylab("UID") +
    theme(axis.text.y = element_text(size=6)) +
    coord_cartesian(xlim = c(-40, 50)) +
    annotate("text", x=-35, y="900.1", label="B")

ACplot <- ggplot(dat, aes(y=factor(UID))) + 
    geom_segment(aes(x=ELnew, xend=ERnew, yend=factor(UID)), 
                 color="blue", size=2) +
    geom_segment(aes(x=ACLnew, xend=ACRnew, yend=factor(UID)), 
                     color="red", size=2, alpha=.5) +
    #ggtitle("Antibody data") + 
    xlab("time since last possible exposure") +
    ylab("UID") +
    theme(axis.text.y = element_text(size=6)) +
    coord_cartesian(xlim = c(-40, 50)) +
    annotate("text", x=-35, y="900.1", label="C")

grid.arrange(IPplot, VCplot, ACplot, ncol=1)
```

**Figure S1:** The intervals observed for each indivdual. Blue intervals represent windows of possible exposure. Red intervals represent windows of possible time of **(A)** symptom onset, **(B)** viral clearance, or **(C)** seroconversion. Lack of a red interval indicates no data for that outcome for that observation.


## 1.2 Bayesian MCMC Framework for Estimating Key Distributions

Statistical analysis roughly follows methods previously used to estimate the incubation period (see @reich2009, @lessler2009). Using the JAGS program and the R statstical language, we implemented and fit the statistical framwork described below.

We treat the time of infection for each individual, $E_i$ as drawn from a uniform prior defined by the earliest and lates possible times of exposure ($EL_i$ and $ER_i$).
$$E_i \sim Uniform(EL_i, ER_i)$$

We then considered the length of the incubation period, $Y_{si}$, the time until viral clearance, $Y_{vi}$, and the time until seroconversion, $Y_{ai}$, to be interval censored random variables following a lognormal, Weibull and Weibull distribution respectively. That is:
$$Y_{si}|E_i \in [SL_i-E_i, SR_i-E_i]; Y_{si} \sim LogNormal(lm, lsd)$$
$$Y_{vi}|E_i \in [VL_i-E_i, VR_i-E_i]; Y_{vi} \sim Weibull(shape_v, scale_v)$$
$$Y_{ai}|E_i \in [AL_i-E_i, AR_i-E_i]; Y_{ai} \sim Weibull(shape_a, scale_a)$$

where $SL_i$ is the earliest possible time of symptom onset for case $i$, $SR_i$ is the latest time of symptom onset for case $i$, $VL_i$ and $VR_i$ similarly bounds viral clearance, and $AL_i$ and $AR_i$ bounds seroconversion. Any censoring bound that was less than 0 was replaced by a very small postive value $\epsilon$.

In our results, the Weibull distribution is parameterized as in the R Statisistical Language, that is, if $a$=shape and $b$=scale, the probability distribution function is defined as:
$$f(x) = \frac{a}{b}\left(\frac{x}{b}\right)^{a-1}e^{-x/b^a}$$

```{r, echo=FALSE, warning=FALSE, results='hide', message=FALSE}

  #assuming observationsiwth no upper limite have NAs in the R slot. Currently only in VCR column
  zika.anal.jags <- zika.anal
  zika.anal.jags$VCR[which(is.infinite(zika.anal.jags$VCR))] <-  NA

  #missing observation are just unbounded, replace with 0
  zika.anal.jags$VCL[which(is.na(zika.anal.jags$VCL))] <-  0
  zika.anal.jags$ACL[which(is.na(zika.anal.jags$ACL))] <-  0


  #make a data object for JAGS
  jags.data <- list(       
                  ER=zika.anal.jags$ER,
                  SL=zika.anal.jags$SL,
                  SR=zika.anal.jags$SR,
                  VL=zika.anal.jags$VCL,
                  VR=zika.anal.jags$VCR, 
                  AL=zika.anal.jags$ACL,
                  AR=zika.anal.jags$ACR)


  #Let jags know the censoring situation
  #1 means interval censored
  #2 means event occured after known time
  jags.data$VSPisCensored=rep(1,25)
  jags.data$VSPisCensored[which(is.na(jags.data$VR))]=2

  jags.data$IPisCensored=rep(1,25)
  jags.data$IPisCensored[which(is.na(jags.data$SR))]=2

  jags.data$ASPisCensored=rep(1,25)
  jags.data$ASPisCensored[which(is.na(jags.data$AR))]=2

  require(rjags)#installed JAGS must be 4.1 or higher

  #define variablees to hold the length of the
  #time to event for symptoms, viral clearance, and seroconversion
  jags.data$Y_S <- rep(NA, 25)
  jags.data$Y_V <- rep(NA, 25)
  jags.data$Y_A <- rep(NA,25)

  #set the intitial values for time to event (i.e., Y_S, Y_V and Y_A)
  IPyInit =  jags.data$SL
  IPyInit[which(is.na(jags.data$SL)==T)]=0
  IPyInit[which(IPyInit==0)]=0.0000000011

  ASPyInit =  jags.data$AL
  ASPyInit[which(is.na(jags.data$AL)==T)]=0
  ASPyInit[which(ASPyInit==0)]=0.0000000011


  VSPyInit =  jags.data$VL
  VSPyInit[which(is.na(jags.data$VL)==T)]=0
  VSPyInit[which(VSPyInit==0)]=0.0000000011
  

  #identify whcih observations are interval censored and which are rihg censored. 
  jags.data$dic=which(jags.data$IPisCensored==1) #this is all inclubation period observtions
  jags.data$Adic=which(jags.data$ASPisCensored==1) 
  jags.data$Arc=which(jags.data$ASPisCensored==2) 
  jags.data$Vdic=which(jags.data$VSPisCensored==1)
  jags.data$Vrc=which(jags.data$VSPisCensored==2)
  
 
  
  #set the parameters we want to track
  parameters <- c("lm","lsd","v_v","scale_v","v_a","scale_a", "E")

  set.seed(12345) #if this is not included, multiple initializations maybe needed.
  #initialization function for jags
  jags.inits <-  function() {
    rc <-list(E=rep(0.0000000011,25), #start with a fixed E to avoid bad starting points
              lm=runif(1,log(2),log(10)),
              lsd = runif(1,.1,log(3)), 
              v_a=runif(1,1,10),
              scale_a= runif(1,0,1), 
              v_v = runif(1,1,10),
              scale_v = runif(1,0,1), 
              Y_S=IPyInit,
              Y_V=VSPyInit,
              Y_A=ASPyInit)
    print(rc)
    return(rc)
  }
  
```

```{r, eval=FALSE, echo=FALSE}
#not evaluated during knitting, run to get the next section to work
#note, we deal with the burnin manually later.

#initialize JAGS model
jagsfit.LWW <- jags.model(file='DistributionFitLWW.jags', 
                          data=jags.data, 
                          inits=jags.inits, 
                          n.chains=3, quiet=F, 
                          n.adapt=10000)


iters<- 1000000
thin <- 50

full.fit.LWW <- coda.samples(jagsfit.LWW, parameters, n.iter=iters, thin=thin, n.chains=3)

#make all of the chains a single matrix with a burnin removed
ABC1=as.matrix(full.fit.LWW[[1]][,])
ABC2=as.matrix(full.fit.LWW[[2]][,])
ABC3=as.matrix(full.fit.LWW[[3]][,])

ABC1=ABC1[5001:(iters/thin),]
ABC2=ABC2[5001:(iters/thin),]
ABC3=ABC3[5001:(iters/thin),]

#recreate MCMC object for diagnostics
full.fit.LWW<-list(as.mcmc(ABC1), as.mcmc(ABC2), as.mcmc(ABC3)) 
chains.LWW <- rbind(ABC1,ABC2,ABC3)
colnames(chains.LWW) <- varnames(full.fit.LWW[[1]])
chains.LWW <- as.data.frame(chains.LWW)
save(full.fit.LWW, chains.LWW, file="../full.fit.jags.LWW.RData")
```


We ran three serperate chains for 1,000,000 iterations each, thinning to 1 in 50 samples. The first 5,000 of the 20,000 retained samples were discarded as burnin. Convergance was assessed by visual inspection and Gelman and Rubin's $\hat{R}$ statistic as well as the multivariate equivalent by Gelman and Brooks [@brooks1998].

**Table S1:** $\hat{R}$ statistic for primary results.
```{r, echo=FALSE}
load("../full.fit.jags.LWW.RData")
print(gelman.diag(full.fit.LWW))
```

# 2. Estimates of key distributions.


**Table S2**: Distribution of parameters and key quantiles of the incubation period for Zika virus infection.

```{r, results='asis', echo=FALSE, message=FALSE}
require(knitr)
inc.fit.jags.LWW <-  quantile(exp(chains.LWW$lm+chains.LWW$lsd^2/2), prob=c(0.5,0.025,0.975))
inc.fit.jags.LWW <- rbind(inc.fit.jags.LWW, 
                         exp(c(median(chains.LWW$lm), quantile(chains.LWW$lm,prob=c(0.025,0.975)))))
inc.fit.jags.LWW <- rbind(inc.fit.jags.LWW,
        exp(c(mean(chains.LWW$lsd), quantile(chains.LWW$lsd,prob=c(0.025,0.975)))))
for (q in c(0.05, 0.25, 0.5, 0.75, 0.95)) {
  tmp <- qlnorm(q, chains.LWW$lm, chains.LWW$lsd)
  inc.fit.jags.LWW <- rbind(inc.fit.jags.LWW,
                        c(mean(tmp), quantile(tmp, prob=c(0.025, 0.975))))
}


colnames(inc.fit.jags.LWW) <- c("est","CIlow","CIhigh")
rownames(inc.fit.jags.LWW) <- c("mean","median",
                            "dispersion",
                            "p5","p25","p50","p75","p95")


kable(inc.fit.jags.LWW, format="markdown", digits=2)

```


**Table S3**: Weibull distribution parameters and key quantiles of the distribution of time to viral clearance for Zika virus infection.

```{r, results='asis', echo=FALSE, message=FALSE}
#in stan alpha = shape in R and beta = rate in R
vshed.fit.jags.LWW <- c(median(chains.LWW$v_v), 
                    quantile(chains.LWW$v_v,prob=c(0.025,0.975)))
vshed.fit.jags.LWW <- rbind(vshed.fit.jags.LWW,
        c(median(chains.LWW$scale_v), 
          quantile(chains.LWW$scale_v,prob=c(0.025,0.975))))
for (q in c(0.05, 0.25, 0.5, 0.75, 0.95, 0.99)) {
  tmp <- qweibull(q, chains.LWW$v_v, chains.LWW$scale_v)
  vshed.fit.jags.LWW <- rbind(vshed.fit.jags.LWW,
                        c(median(tmp), quantile(tmp, prob=c(0.025, 0.975))))
}


tmp <- chains.LWW$scale_v * gamma(1+1/chains.LWW$v_v)
vshed.fit.jags.LWW <- rbind(quantile(tmp,prob=c(.5,.025,.975)),
                           vshed.fit.jags.LWW)
colnames(vshed.fit.jags.LWW) <- c("est","CIlow","CIhigh")
rownames(vshed.fit.jags.LWW) <- c("mean",
                            "shape",
                            "scale",
                            "p5","p25","p50","p75","p95","p99")



kable(vshed.fit.jags.LWW, format="markdown", digits=2)

```



**Table S4**: Weibull distribution parameters and key quantiles of the distribution of time to seroconversion for Zika virus infection.

```{r, results='asis', echo=FALSE, message=FALSE}
#in stan alpha = shape in R and beta = rate in R
sero.fit.jags.LWW <- c(median(chains.LWW$v_a), 
                    quantile(chains.LWW$v_a,prob=c(0.025,0.975)))
sero.fit.jags.LWW <- rbind(sero.fit.jags.LWW,
        c(median(chains.LWW$scale_a), 
          quantile(chains.LWW$scale_a,prob=c(0.025,0.975))))
for (q in c(0.05, 0.25, 0.5, 0.75, 0.95, 0.99)) {
  tmp <- qweibull(q, chains.LWW$v_a, chains.LWW$scale_a)
  sero.fit.jags.LWW <- rbind(sero.fit.jags.LWW,
                        c(median(tmp), quantile(tmp, prob=c(0.025, 0.975))))
}


tmp <- chains.LWW$scale_a * gamma(1+1/chains.LWW$v_a)
sero.fit.jags.LWW <- rbind(quantile(tmp,prob=c(.5,.025,.975)),
                           sero.fit.jags.LWW)
colnames(sero.fit.jags.LWW) <- c("est","CIlow","CIhigh")
rownames(sero.fit.jags.LWW) <- c("mean",
                                 "shape",
                                "scale",
                              "p5","p25","p50","p75","p95","p99")


kable(sero.fit.jags.LWW, format="markdown", digits=2)


```


```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=2.5, fig.height=5}
#Try to use ggplot to plot a beautiful display of the data. 

#first make data frames to hold everything
inc.curve <- NULL
vc.curve <- NULL
sero.curve<- NULL

for (q in seq(0,30,.1)) {
    tmp <- plnorm(q, chains.LWW$lm, chains.LWW$lsd)
    tmp <- quantile(tmp, prob=c(0.025, .5, 0.975))
    inc.curve <- rbind(inc.curve, c(q=q,
                                    plow=tmp[1],
                                    pmid=tmp[2],
                                    phigh=tmp[3]))
    
    tmp <- 1-pweibull(q, chains.LWW$v_v, chains.LWW$scale_v)
    tmp <- quantile(tmp, prob=c(0.025, .5, 0.975))
    vc.curve <- rbind(vc.curve, c(q=q,
                                  plow=tmp[1],
                                  pmid=tmp[2],
                                  phigh=tmp[3]))
    
    tmp <- pweibull(q, chains.LWW$v_a, chains.LWW$scale_a)
    tmp <- quantile(tmp, prob=c(0.025, .5, 0.975))
    sero.curve <- rbind(sero.curve, c(q=q,
                                  plow=tmp[1],
                                  pmid=tmp[2],
                                  phigh=tmp[3]))
}
inc.curve <- as.data.frame(inc.curve)
colnames(inc.curve) <- c("q","incplow","incpmid","incphigh")

vc.curve <- as.data.frame(vc.curve)
colnames(vc.curve) <- c("q","vcplow","vcpmid","vcphigh")

sero.curve <- as.data.frame(sero.curve)
colnames(sero.curve) <- c("q","scplow","scpmid","scphigh")


#require(ggplot2)


inc.plt <- ggplot(inc.curve, aes(x=q)) +
  geom_ribbon(aes(ymin=incplow, ymax=incphigh), fill="blue", alpha=.2) +
  geom_line(aes(y=incpmid), col="blue") +theme_bw()+
  scale_x_continuous(limits=c(0, 25), expand = c(0, 0)) + 
  ylab("% w/ symptoms") + xlab("days")

vc.plt <- ggplot(vc.curve, aes(x=q)) +
  geom_ribbon(aes(ymin=vcplow, ymax=vcphigh), fill="red", alpha=.2) +
  geom_line(aes(y=vcpmid), col="red") +theme_bw()+
  scale_x_continuous(limits=c(0, 25), expand = c(0, 0)) + 
  ylab("% w/ detectable virus") + xlab("days")

sero.plt <- ggplot(sero.curve, aes(x=q)) +
  geom_ribbon(aes(ymin=scplow, ymax=scphigh), fill="darkgreen", alpha=.2) +
  geom_line(aes(y=scpmid), col="darkgreen")+theme_bw()+
  scale_x_continuous(limits=c(0, 25), expand = c(0, 0)) + 
   ylab("% w/ sero-positive") + xlab("days")


#require(gridExtra)
grid.arrange(inc.plt, vc.plt, sero.plt, ncol=1)

```

**Figure S2:** Percent of the population who eventually develop symptoms who do so by day t (top), percent with detectable virus in blood at day t (middle), and percent seroconverting by day t (bottom).


# 3. Blood Supply Safety
 
For a fixed daily incidence rate of Zika virus infection, $\lambda$, the expected probability of a person donating infected blood if there was no screening (this includes allowing symptomatic people to donate blood) is:
$$\mu_v \lambda$$
where $\mu_v$ is the mean time until viral clearance and we assume that blood donations would be infectious from the moment a case is infected.

We presume that 80% of those infected to Zika virus remain asymptomatic, and that those who do develop symptoms always do so before clearing the virus. Hence the reduction in the probablity of an infected blood donation that can be achieved by preventing those who currently or recently have had symptoms associated with Zika virus from donating blood is:
$$\frac{0.8\mu_v + 0.2 \mu_s}{\mu_v}$$
where $\mu_s$ is the mean incubation period. 

To estimate the effect of a serological screening test, we make the simplifying assumption that the time to seroconversion and the time to viral clearance are independent, and calculate the mean time to clearing virus or seroconverting, $\mu_{vs}$, based on the area under the survival curves as:
$$\mu_{vs} = \int_0^\infty (1-F(t))(1-G(t)) dt$$
where $F(t)$ is the CDF for the Weibull distribution of time for viral clearance, and $G(t)$ is the CDF for the Weibull distribution for time to seroconversion. Hence the reduction in the probability of a infected blood donation by serological screening is:
$$\frac{\mu_{vs}}{\mu_v}$$


We applied these calculations to our derived distributions, calculating point estimates and 95% confidence intervals by calculating the estimate for each  sample from the posterior of the MCMC fit of key distributions, and taking the median and appropriate quantiles.

**Table S5:**  Increase in risk of a Zika contaminated donation for each 1 per 100,000 
person-day increase in Zika incidence.
```{r, results='asis', echo=FALSE}

#caclculate the virologic and symptom means and compare.
mns <- data.frame(inc = 0.2*exp(chains.LWW$lm + chains.LWW$lsd^2/2)+0.8* chains.LWW$scale_v*gamma(1+1/chains.LWW$v_v),
                  viro = chains.LWW$scale_v*gamma(1+1/chains.LWW$v_v))

relative.inc <- quantile(mns$inc/mns$viro, 
                              prob=c(0.025, .5, 0.975))


to.integrate <- function(t, v_v, scale_v, v_a, scale_a) {
   rc <- (1-pweibull(t,v_v, scale_v)) *
      (1-pweibull(t,v_a, scale_a))
   return(rc)
  }


#mean survival time for serologic and virologic. 
mns$sero.viro <- NA
for (i in 1:length(chains.LWW$lm)) {
  mns$sero.viro[i] <- 
    integrate(to.integrate, 0,Inf, 
              v_v= chains.LWW$v_v[i],
              scale_v= chains.LWW$scale_v[i],
              v_a= chains.LWW$v_a[i],
              scale_a= chains.LWW$scale_a[i])$value
  }

relative.sero <- quantile(mns$sero.viro/mns$viro,
                                  prob=c(0.025, .5, 0.975))


tmp <- apply(mns, 2, quantile, prob=c(0.5, 0.025, 0.975))
tmp <- tmp[,c("viro","inc","sero.viro")]
screen.res <-data.frame(risk.per.100k=sprintf("%1.1f (%1.1f, %1.1f)",
                                 tmp[1,],tmp[2,], tmp[3,]),
           relativerisk=c("---",
                           sprintf("%1.2f (%1.2f, %1.2f)", 
                                    relative.inc[2], relative.inc[1], relative.inc[3]),
                           sprintf("%1.2f (%1.2f, %1.2f)", 
                                    relative.sero[2], relative.sero[1], relative.sero[3]))                    
                          
             )

colnames(screen.res) <- c("Risk per 100,000 donations", "Relative Risk")
rownames(screen.res) <- c("No Screening", "Symptom Based Screening", "Serological Screening")

kable(screen.res, format="markdown", digits=2)

```

#4. Distribution of times to virus recovery in other specimens

We were unable to measure the distribution of time to virus no longer being recoverable in samples other than blood/sera (e.g., urine, saliva) due to insufficient data. However, we can examine the posterior distribution of the time from infection to collection of postive samples. We confine our analysis to urine samples because of the concern for sexual transmission, and because urine samples comprise most of the reported non-blood tests.


**Table S6:** Expected time from infection to collection of latest negative urine sample.
```{r, echo=FALSE, results='asis'}
zika.urine<- zika[which(zika$SAMPLE=="urine" & zika$SHEDDING==1),]

#recenter VL to 0 from EL
zika.urine$VL  <- zika.urine$VL - zika.urine$EL

#simplify the data to just what we need...all of these war within a day
zika.urine <- zika.urine[,c("UID","VL")]

#make sure we have only the latest sample for each individual
for (uid in unique(zika.urine$UID)) {
  inds <- which(zika.urine$UID == uid)
  zika.urine$VL[inds] <- max(zika.urine$VL[inds])
}

zika.urine <- unique(zika.urine)

zika.urine$anal.ind <- -1
zika.urine$TimeToTest <- -1

#get the index in the analysis indices and calculate the time from exposure
for (i in 1:nrow(zika.urine)) {
  zika.urine$anal.ind[i] <- which(zika.anal$UID==zika.urine$UID[i])
  tmp <- quantile(zika.urine$VL[i]-chains.LWW[,zika.urine$anal.ind[i]], prob=c(0.5,0.025,0.975))
  zika.urine$TimeToTest[i] <- sprintf("%1.1f (%1.1f, %1.1f)",
                                   tmp[1],tmp[2],tmp[3])
}

kable(zika.urine[,c("UID","TimeToTest")], format="markdown", digits=1, row.names=FALSE)
```

# 5. Model Checking and Sensitivity Analysis

We examined several other distributions and mechanisms for fitting models. Many (e.g., those assuming a gamma distribution for the viral clearance an seroconversion distributions) could not be fit. Here we compare the results from the most similar plausible model, assuming a log normal distribution for all times to events, with our main results. Note that the distributions used in the main analysis were selected a priori to be log-normal for the incubation period, and gamma or Weibull for seroconversion and viral shedding (the former of which could not be successfully fit). Due to limited data and lack of qualitative differences, no formal comparison of fit was performed.

## Use of an Alternate Distribution
```{r, eval=FALSE, echo=FALSE}
 #code for running the LLL model

#LLL model
#not evaluated during knitting, run to get the next section to work
#note, we deal with the burnin manually later.


  #set the parameters we want to track
  parameters <- c("lm","lsd","V_lm","V_lsd","A_lm","A_lsd", "E")


  set.seed(275781116)
  #initialization function for jags
  jags.inits <-  function() {
    rc <-list(E=rep(0.0000000011,25), #start with a fixed E to avoid bad starting points
              lm=runif(1,log(2),log(10)),
              lsd = runif(1,.1,log(3)), 
              A_lm=runif(1,log(2),log(10)),
              A_lsd = runif(1,.1,log(3)), 
              V_lm=runif(1,log(2),log(10)),
              V_lsd = runif(1,.1,log(3)), 
              Y_S=IPyInit,
              Y_V=VSPyInit,
              Y_A=ASPyInit)
    print(rc)
    return(rc)
  }
  

  

#initialize JAGS model
jagsfit.LLL <- jags.model(file='DistributionFitLLL.jags', 
                          data=jags.data, 
                          inits=jags.inits, 
                          n.chains=3, quiet=F, 
                          n.adapt=10000)


iters<- 1000000
thin <- 50

full.fit.LLL <- coda.samples(jagsfit.LLL, parameters, n.iter=iters, thin=thin, n.chains=3)

#make all of the chains a single matrix with a burnin removed
ABC1=as.matrix(full.fit.LLL[[1]][,])
ABC2=as.matrix(full.fit.LLL[[2]][,])
ABC3=as.matrix(full.fit.LLL[[3]][,])

ABC1=ABC1[5001:(iters/thin),]
ABC2=ABC2[5001:(iters/thin),]
ABC3=ABC3[5001:(iters/thin),]

full.fit.LLL <- rbind(ABC1,ABC2,ABC3)


#recreate MCMC object for diagnostics
full.fit.LLL<-list(as.mcmc(ABC1), as.mcmc(ABC2), as.mcmc(ABC3)) 
chains.LLL <- rbind(ABC1,ABC2,ABC3)
colnames(chains.LLL) <- varnames(full.fit.LLL[[1]])
chains.LLL <- as.data.frame(chains.LLL)
save(full.fit.LLL, chains.LLL, file="../full.fit.jags.LLL.RData")


```


**Table S7**: Distribution of parameters and key quantiles of the incubation period for Zika virus infection when assuming all distributions are lognormal. 

```{r, results='asis', echo=FALSE, message=FALSE}
load("../full.fit.jags.LLL.RData")
require(knitr)
inc.fit.jags.LLL <-  quantile(exp(chains.LLL$lm+chains.LLL$lsd^2/2), prob=c(0.5,0.025,0.975))
inc.fit.jags.LLL <- rbind(inc.fit.jags.LLL,exp(c(median(chains.LLL$lm), quantile(chains.LLL$lm,prob=c(0.025,0.975)))))
inc.fit.jags.LLL <- rbind(inc.fit.jags.LLL,
        exp(c(mean(chains.LLL$lsd), quantile(chains.LLL$lsd,prob=c(0.025,0.975)))))
for (q in c(0.05, 0.25, 0.5, 0.75, 0.95)) {
  tmp <- qlnorm(q, chains.LLL$lm, chains.LLL$lsd)
  inc.fit.jags.LLL <- rbind(inc.fit.jags.LLL,
                        c(mean(tmp), quantile(tmp, prob=c(0.025, 0.975))))
}


colnames(inc.fit.jags.LLL) <- c("est","CIlow","CIhigh")
rownames(inc.fit.jags.LLL) <- c("mean",
                            "median",
                            "dispersion",
                            "p5","p25","p50","p75","p95")


kable(inc.fit.jags.LLL, format="markdown", digits=2)

```

**Table S8**: Lognormal distribution parameters and key quantiles of the distribution of time to viral clearance for Zika virus infection assuming all distributions are lognormal.


```{r, results='asis', echo=FALSE, message=FALSE}

require(knitr)
vshed.fit.jags.LLL <-  quantile(exp(chains.LLL$V_lm+chains.LLL$V_lsd^2/2), prob=c(0.5,0.025,0.975))
vshed.fit.jags.LLL <- rbind(vshed.fit.jags.LLL,exp(c(median(chains.LLL$V_lm), quantile(chains.LLL$V_lm,prob=c(0.025,0.975)))))
vshed.fit.jags.LLL <- rbind(vshed.fit.jags.LLL,
        exp(c(mean(chains.LLL$V_lsd), quantile(chains.LLL$V_lsd,prob=c(0.025,0.975)))))
for (q in c(0.05, 0.25, 0.5, 0.75, 0.95,0.99)) {
  tmp <- qlnorm(q, chains.LLL$V_lm, chains.LLL$V_lsd)
  vshed.fit.jags.LLL <- rbind(vshed.fit.jags.LLL,
                        c(mean(tmp), quantile(tmp, prob=c(0.025, 0.975))))
}


colnames(vshed.fit.jags.LLL) <- c("est","CIlow","CIhigh")
rownames(vshed.fit.jags.LLL) <- c("mean",
                            "median",
                            "dispersion",
                            "p5","p25","p50","p75","p95","p99")


kable(vshed.fit.jags.LLL, format="markdown", digits=2)

```



**Table S9**: Lognormal distribution parameters and key quantiles of the distribution of time to seroconversion for Zika virus infection assuming all distributions are lognormal.


```{r, results='asis', echo=FALSE, message=FALSE}

require(knitr)
sero.fit.jags.LLL <-  quantile(exp(chains.LLL$A_lm+chains.LLL$A_lsd^2/2), prob=c(0.5,0.025,0.975))
sero.fit.jags.LLL <- rbind(sero.fit.jags.LLL,exp(c(median(chains.LLL$A_lm), quantile(chains.LLL$A_lm,prob=c(0.025,0.975)))))
sero.fit.jags.LLL <- rbind(sero.fit.jags.LLL,
        exp(c(mean(chains.LLL$A_lsd), quantile(chains.LLL$A_lsd,prob=c(0.025,0.975)))))
for (q in c(0.05, 0.25, 0.5, 0.75, 0.95,0.99)) {
  tmp <- qlnorm(q, chains.LLL$A_lm, chains.LLL$A_lsd)
  sero.fit.jags.LLL <- rbind(sero.fit.jags.LLL,
                        c(mean(tmp), quantile(tmp, prob=c(0.025, 0.975))))
}


colnames(sero.fit.jags.LLL) <- c("est","CIlow","CIhigh")
rownames(sero.fit.jags.LLL) <- c("mean",
                            "median",
                            "dispersion",
                            "p5","p25","p50","p75","p95","p99")


kable(sero.fit.jags.LLL, format="markdown", digits=2)

```


**Table S10:**  Increase in risk of a Zika contaminated donation for each 1 per 100,000 
person-day increase in Zika incidence, assuming all lognormal distributions
```{r, results='asis', echo=FALSE}

#caclculate the virologic and symptom means and compare.
mns <- data.frame(inc = exp(chains.LLL$lm + chains.LLL$lsd^2/2)*.2 + exp(chains.LLL$V_lm + chains.LLL$V_lsd^2/2)*.8,
                  viro = exp(chains.LLL$V_lm + chains.LLL$V_lsd^2/2))

relative.inc <- quantile(mns$inc/mns$viro, 
                              prob=c(0.025, .5, 0.975))


to.integrate <- function(t, V_lm, V_lsd, A_lm, A_lsd) {
   rc <- (1-plnorm(t,V_lm, V_lsd)) *
      (1-plnorm(t,A_lm, A_lsd))
   return(rc)
  }


#mean survival time for serologic and virologic. 
mns$sero.viro <- NA
for (i in 1:length(chains.LLL$lm)) {
  mns$sero.viro[i] <- 
    integrate(to.integrate, 0,Inf, 
              V_lm= chains.LLL$V_lm[i],
              V_lsd= chains.LLL$V_lsd[i],
              A_lm= chains.LLL$A_lm[i],
              A_lsd= chains.LLL$A_lsd[i])$value
  }

relative.sero <- quantile(mns$sero.viro/mns$viro,
                                  prob=c(0.025, .5, 0.975))


tmp <- apply(mns, 2, quantile, prob=c(0.5, 0.025, 0.975))
tmp <- tmp[,c("viro","inc","sero.viro")]
screen.res <-data.frame(risk.per.100k=sprintf("%1.1f (%1.1f, %1.1f)",
                                 tmp[1,],tmp[2,], tmp[3,]),
           relativerisk=c("---",
                           sprintf("%1.2f (%1.2f, %1.2f)", 
                                    relative.inc[2], relative.inc[1], relative.inc[3]),
                           sprintf("%1.2f (%1.2f, %1.2f)", 
                                    relative.sero[2], relative.sero[1], relative.sero[3]))                    
                          
             )

colnames(screen.res) <- c("Risk per 100,000 donations", "Relative Risk")
rownames(screen.res) <- c("No Screening", "Symptom Based Screening", "Serological Screening")

kable(screen.res, format="markdown", digits=2)

```

## Alternate Fitting Mechanisms

Alternate methods of fitting the model and distributions and a spreadsheet comparing methods are available on the GitHub repository. Key quantiles were robust across all methods and assumptions, and the viral shedding distribution was the most sensitive to the selection of the fitting approach.

# References
