---
title: "Exploratory Zika Analysis"
author: "Justin Lessler"
date: "Monday, February 22, 2016"
output: word_document
---

**DISCLAMER: This Exploratory original analysis is provided in the repository for completeness. There is no guarantee it will work with the final analytic data set, and it contains many intermediate analyses that did not pan out, of for which we did not ultimately have condidence in the results. It also will not knot efficiently.**

Pooled analysis of data form the Zika systematic review.

First load the data and figure out how big our data set is for each purpose.

```{r}
zika <- read.csv("ZikaLitReviewData.csv")


#DROP PERINATAL TRANSMISSIONS
zika <- zika[which(zika$TRANS.MECH!="perinatal"),]  

#How many people do we have?
length(unique(zika$UID))

#Now lets divide data sets by observation type.

#Incubation period
zika.inc <- zika[!is.na(zika$SL),]
nrow(zika.inc)

#Period of viral shedding
zika.vshed <- zika[!is.na(zika$VL),]
nrow(zika.vshed)
length(unique(zika.vshed$UID))

#Time to seropositiveity
zika.sero <- zika[!is.na(zika$AL),]
nrow(zika.sero)
length(unique(zika.sero$UID))

```

So we have data with which to estimate the incubation period from 25 people; 46 observations from 20 people to estimate the length of viral shedding; and 63 observations from 22 people to estimate the time to seropositivity.


# coarseDataTools Bootstrap based analyses

We first estimate each relative distribution using coarseDataTools and bootstrapping. Replicating main analysis, assuming a log-normal distribution for the incubation period, and Weibull for the other relevant distributions. 


## Incubation Period

Basic incubation period analysis assuming a log-normal incubation period and using a bootstrap method for calculating confidence intervals. 

```{r, results='hide'}
library(coarseDataTools)

#make the data to pass in to dic.fit
inc.dat <- zika.inc[,c("EL","ER","SL","SR")]
inc.dat$type <- 0

#fit it...use bootstrap interstions for now. MCMC better?
inc.fit <- dic.fit(inc.dat,dist="L",n.boots=5000,
                   ptiles = c(0.05, 0.25, 0.5, 0.75, 0.95))

```

```{r}
plot(inc.fit, xlim=c(0,20), plot.n.samps=2000)
inc.fit
```



##Length of viral shedding



```{r}

#first pass, restrict to blood (i.e., not saliva or uring)
zika.vshed.bl <- zika.vshed[which(zika.vshed$SAMPLE.TYPE %in%c("serum","blood")),]

#Go through and find the start of the latest period in which
#someone tested seropositive and the end of the latest subsequent
#period where they did not....these bound seroconversion
for (uid in unique(zika.vshed.bl$UID)) {
  inds <- which(zika.vshed.bl$UID==uid)
  tmp <- zika.vshed.bl[inds,] 
  
  #taking greater than rqualt to 1 means we will treat equivocal results as 
  #positive
  low.time <- max(tmp$VL[tmp$SHEDDING>=1])
  
  #no earliest point...go to earliest time of infection
  if (low.time<tmp$EL[1]) {low.time <- tmp$EL[1]}
  
  high.time <- min(tmp$VR[tmp$SHEDDING==0 &  tmp$VR>low.time])
  ##print(tmp) ###DEBUG
  ##cat(uid, low.time, high.time,"\n") ###DEBUG
  
  zika.vshed.bl$VCL[inds] <- low.time
  zika.vshed.bl$VCR[inds] <- high.time
  }


  #make the anlytic dataset
  vshed.bl.anal <- zika.vshed.bl[,c("UID","EL","ER","VCL","VCR")]
  vshed.bl.anal$type <- 0
  vshed.bl.anal <- unique(vshed.bl.anal) #one row per observation

  ###SUPECEDE BY JOINT STAN ANALYSIS, BUT KEPT FOR COMPARISON
  tmp <- vshed.bl.anal 
  colnames(tmp) <- c("UID","EL","ER","SL","SR","type")
  tmp$SL <- tmp$SL+.01 #avoids some problems with 0s

 #change the infinite SRs to something really big
   tmp$SR[is.infinite(tmp$SR)] <- 
       tmp$ER[is.infinite(tmp$SR)] + 730
 

  ####A BIT OF A DISASTER
  #now lets fit a gamma distribution to this data using MCMC
  vshed.fit <- dic.fit(tmp, dist = "G", n.boots = 10000)
  

```


# Time to Seroconversion
```{r}
# For the first pass we will assume any seropositivity indicates
#serpositivity by IgG

zika.sero$ACL <- NA
zika.sero$ACR <- NA

#for the purposes of this analysis assume that a
#NA in any antibody should be replaces with the IgM value
zika.sero$ANYANTIBODY[is.na(zika.sero$ANYANTIBODY)] <- 
  zika.sero$IGM[is.na(zika.sero$ANYANTIBODY)]


#Loop through unique IDs bounding their time of 
#seroconversion
for (uid in unique(zika.sero$UID)) {
  inds <- which(zika.sero$UID == uid)
  tmp <- zika.sero[inds,]
  
  
  # print(inds) # DEBUG
  
  #lowest time is the latest time with a negative serology
  low.time <- max(tmp$AL[tmp$ANYANTIBODY==0])
  
   
  #no earliest point...go to earliest possible  time of infection
  if (low.time<tmp$EL[1]) {low.time <- tmp$EL[1]}
  
  #latest possible time is the earliest time of a positive antibody test. For our
  #purposes, assume that equivocal is seropositive
  high.time <- min(tmp$AR[which(tmp$ANYANTIBODY>0)])
  
  
  zika.sero$ACL[inds] <- low.time
  zika.sero$ACR[inds] <- high.time
  
}

#make an analytic dataset with one row per person
sero.anal <- zika.sero[,c("UID", "EL", "ER", "ACL","ACR")]
sero.anal$type <- 0 
sero.anal <- unique(sero.anal)


#Fit using coarse data tools methods
#for the most part this should be considered to be superceded by RStan stuff

############REJECTS A LOT OF BOOTSTRAPS, BUT WORKS

tmp <- sero.anal
colnames(tmp) <- c("UID","EL","ER","SL","SR","type")

tmp$SR[is.infinite(tmp$SR)] <- 
  tmp$ER[is.infinite(tmp$SR)] + 730

sero.fit<- dic.fit(tmp, dist = "W", n.boots = 10000)


```



#Combined stan analysis Lognomal for incubation period, Weibull for others

```{r, eval=FALSE}
require(rstan)


#Create a dataset with everythign lined up for passing into STAN
zika.anal <- zika.inc[,c("UID","EL","ER","SL","SR")]
zika.anal <- merge(zika.anal, vshed.bl.anal, all=TRUE)
zika.anal$type <- 0 #we don't use this, but helps with the merge.
zika.anal <- merge(zika.anal, sero.anal, all=TRUE)

zika.anal$SL[which(is.na(zika.anal$SL))] <- 
  zika.anal$EL[which(is.na(zika.anal$SL))]
zika.anal$SR[which(is.na(zika.anal$SR))] <- 
   zika.anal$ER[which(is.na(zika.anal$SR))] + 100

zika.anal$ACL[which(is.na(zika.anal$ACL))] <-
  zika.anal$EL[which(is.na(zika.anal$ACL))]
zika.anal$ACR[which(is.na(zika.anal$ACR))] <- 
  zika.anal$ER[which(is.na(zika.anal$ACR))] + 720

zika.anal$VCL[which(is.na(zika.anal$VCL))] <-
  zika.anal$EL[which(is.na(zika.anal$VCL))]
zika.anal$VCR[which(is.na(zika.anal$VCR))] <- 
  zika.anal$ER[which(is.na(zika.anal$VCR))] + 720


zika.anal$VCR[which(is.infinite(zika.anal$VCR))] <- 
  zika.anal$ER[which(is.infinite(zika.anal$VCR))] + 720

#renormalize everything so EL is 0
zika.anal[,c("EL","ER","SL","SR","VCL","VCR","ACL","ACR")] <-
  zika.anal[,c("EL","ER","SL","SR","VCL","VCR","ACL","ACR")] -zika.anal$EL  


#for STAN purposes, it is usefule if all of these variables index 
#from 0
stan.data <- list(N=nrow(zika.anal),
                  EL=zika.anal$EL,
                  ER=zika.anal$ER,
                  SL=zika.anal$SL,
                  SR=zika.anal$SR,
                  VL=zika.anal$VCL,
                  VR=zika.anal$VCR, 
                  AL=zika.anal$ACL,
                  AR=zika.anal$ACR)


#initialization version for the lognormal/weibull stan model


#use uniform from halfway trhough the exposure interval to ER.
#still fails occasionally, but MUCH less often. 
stan.init.LWW <- function() {
    rc <-list(E=runif(stan.data$N, (stan.data$EL+stan.data$ER)/2, stan.data$ER),
                lm= runif(1,log(2),log(10)),
                lsd = runif(1,.1,2),
                v_alpha = 2,#,runif(1,5,10),
                v_sigma= 50,# runif(1,.5,2) 
                s_alpha = 2,#,runif(1,5,10),
                s_sigma= 50# runif(1,.5,2) 
                )
    #print(rc) #DEBUG
    #curve(dweibull(x,rc$v_alpha, rc$v_sigma), from=0, to=50) #DEBUG
    return(rc)
    }



#Need at least 35,000 iterations to converge
full.fit.LWW <- stan("DistributionFitLWW.stan", data=stan.data,
                  init=stan.init.LWW,
                  iter=50000, chains=4, thin=10)

save(full.fit.LWW, file = "../FullFitLWW.RData")

```

```{r}
load(file = "../FullFitLWW.RData")

#now we need to calculate the quantiles from the chains and construct our
#data frame
chains <- extract(full.fit.LWW)


inc.fit.stan.LWW <- c(median(chains$lm), quantile(chains$lm,prob=c(0.025,0.975)))
inc.fit.stan.LWW <- rbind(inc.fit.stan.LWW,
        c(mean(chains$lsd), quantile(chains$lsd,prob=c(0.025,0.975))))
for (q in c(0.05, 0.25, 0.5, 0.75, 0.95)) {
  tmp <- qlnorm(q, chains$lm, chains$lsd)
  inc.fit.stan.LWW <- rbind(inc.fit.stan.LWW,
                        c(mean(tmp), quantile(tmp, prob=c(0.025, 0.975))))
}

colnames(inc.fit.stan.LWW) <- c("est","CIlow","CIhigh")
rownames(inc.fit.stan.LWW) <- c("meanlog",
                            "sdlog",
                            "p5","p25","p50","p75","p95")



#in stan alpha = shape in R and beta = rate in R
vshed.fit.stan.LWW <- c(median(chains$v_alpha), 
                    quantile(chains$v_alpha,prob=c(0.025,0.975)))
vshed.fit.stan.LWW <- rbind(vshed.fit.stan.LWW,
        c(mean(chains$v_sigma), 
          quantile(chains$v_sigma,prob=c(0.025,0.975))))
for (q in c(0.05, 0.25, 0.5, 0.75, 0.95, 0.99)) {
  tmp <- qweibull(q, chains$v_alpha, chains$v_sigma)
  vshed.fit.stan.LWW <- rbind(vshed.fit.stan.LWW,
                        c(mean(tmp), quantile(tmp, prob=c(0.025, 0.975))))
}

colnames(vshed.fit.stan.LWW) <- c("est","CIlow","CIhigh")
rownames(vshed.fit.stan.LWW) <- c("alpha (shape)",
                            "sigma (scale)",
                            "p5","p25","p50","p75","p95","p99")



#sero
sero.fit.stan.LWW <- c(median(chains$s_alpha), 
                    quantile(chains$s_alpha,prob=c(0.025,0.975)))
sero.fit.stan.LWW <- rbind(sero.fit.stan.LWW,
        c(mean(chains$s_sigma), 
          quantile(chains$s_sigma,prob=c(0.025,0.975))))
for (q in c(0.05, 0.25, 0.5, 0.75, 0.95, 0.99)) {
  tmp <- qweibull(q, chains$s_alpha, chains$s_sigma)
  sero.fit.stan.LWW <- rbind(sero.fit.stan.LWW,
                        c(mean(tmp), quantile(tmp, prob=c(0.025, 0.975))))
}

colnames(sero.fit.stan.LWW) <- c("est","CIlow","CIhigh")
rownames(sero.fit.stan.LWW) <- c("alpha (shape)",
                            "sigma (scale)",
                            "p5","p25","p50","p75","p95","p99")

inc.fit.stan.LWW

vshed.fit.stan.LWW
cat("Mean viral shedding length=",
    vshed.fit.stan.LWW[2,1] * gamma(1+1/vshed.fit.stan.LWW[1,1]),
    "(",vshed.fit.stan.LWW[2,2] * gamma(1+1/vshed.fit.stan.LWW[1,2]),
    ",",vshed.fit.stan.LWW[2,3] * gamma(1+1/vshed.fit.stan.LWW[1,3]),
    ")")
                                  

sero.fit.stan.LWW
cat("Mean time to seroconversion=",
    sero.fit.stan.LWW[2,1] * gamma(1+1/sero.fit.stan.LWW[1,1]),
    "(",sero.fit.stan.LWW[2,2] * gamma(1+1/sero.fit.stan.LWW[1,2]),
    ",",sero.fit.stan.LWW[2,3] * gamma(1+1/sero.fit.stan.LWW[1,3]),
    ")")
              
```



Based on 4 chains, 50,000 interations per chain, assuming the incubtion period follows a log-normal distribution, and that the time to viral clearance an seroconversion both follow a weibull distributoin, we find that:

The median incubation period is 5.9 days (95% CI: 4.6-7.5) with a dispersion of 1.45 (95% CI: 1.23-1.91). 90% of cases will develop symptoms between 3.2 and 11.2 days of infection (95% CIs: 1.8-4.6 and 7.9-17.8).

The mean time to viral clearanace is estimated to be 9.8 days (95% CI: 7.7-13.4), and  we expext 90% of people to clear virus between 3.5 and 17.9 days after infection (95% CIs: 0.8-5.6 and 12.5-33.9).

The mean time to seroconversion is 9.1 days (95% CI: 7.0-11.1), and we expect 90% of peopld to seroconvert between 4.0 and 14.3 days of infection (95% CIs: 2.1-5.5 and 11.1-19.1).


```{r}
#Try to use ggplot to plot a beautiful display of the data. 

#first make data frames to hold everything
inc.curve <- NULL
vc.curve <- NULL
sero.curve<- NULL

for (q in seq(0,30,.1)) {
    tmp <- plnorm(q, chains$lm, chains$lsd)
    tmp <- quantile(tmp, prob=c(0.025, .5, 0.975))
    inc.curve <- rbind(inc.curve, c(q=q,
                                    plow=tmp[1],
                                    pmid=tmp[2],
                                    phigh=tmp[3]))
    
    tmp <- 1-pweibull(q, chains$v_alpha, chains$v_sigma)
    tmp <- quantile(tmp, prob=c(0.025, .5, 0.975))
    vc.curve <- rbind(vc.curve, c(q=q,
                                  plow=tmp[1],
                                  pmid=tmp[2],
                                  phigh=tmp[3]))
    
    tmp <- pweibull(q, chains$s_alpha, chains$s_sigma)
    tmp <- quantile(tmp, prob=c(0.025, .5, 0.975))
    sero.curve <- rbind(sero.curve, c(q=q,
                                  plow=tmp[1],
                                  pmid=tmp[2],
                                  phigh=tmp[3]))
}
inc.curve <- as.data.frame(inc.curve)
colnames(inc.curve) <- c("q","incplow","incpmid","incphigh")

vc.curve <- as.data.frame(vc.curve)
colnames(vc.curve) <- c("q","vcplow","vcpmid","vcphigh")

sero.curve <- as.data.frame(sero.curve)
colnames(sero.curve) <- c("q","scplow","scpmid","scphigh")


require(ggplot2)


inc.plt <- ggplot(inc.curve, aes(x=q)) +
  geom_ribbon(aes(ymin=incplow, ymax=incphigh), fill="blue", alpha=.2) +
  geom_line(aes(y=incpmid), col="blue") +theme_bw()+
  scale_x_continuous(limits=c(0, 25), expand = c(0, 0)) + 
  ylab("% w/ symptoms")

vc.plt <- ggplot(vc.curve, aes(x=q)) +
  geom_ribbon(aes(ymin=vcplow, ymax=vcphigh), fill="red", alpha=.2) +
  geom_line(aes(y=vcpmid), col="red") +theme_bw()+
  scale_x_continuous(limits=c(0, 25), expand = c(0, 0)) + 
  ylab("% w/ detectable virus")

sero.plt <- ggplot(sero.curve, aes(x=q)) +
  geom_ribbon(aes(ymin=scplow, ymax=scphigh), fill="darkgreen", alpha=.2) +
  geom_line(aes(y=scpmid), col="darkgreen")+theme_bw()+
  scale_x_continuous(limits=c(0, 25), expand = c(0, 0)) + 
   ylab("% w/ sero-positive")


require(gridExtra)
grid.arrange(inc.plt, vc.plt, sero.plt, ncol=1)

```




Caclcualte the means, which translates into risk.

```{r}

#caclculate the virologic and symptom means and compare.
mns <- data.frame(inc = exp(chains$lm + chains$lsd^2/2),
                  viro = chains$v_sigma*gamma(1+1/chains$v_alpha))

symptom.relative.inc <- quantile((0.8*mns$viro + 0.2*mns$inc)/mns$viro, 
                              prob=c(0.025, .5, 0.975))


to.integrate <- function(t, v_alpha, v_sigma, s_alpha, s_sigma) {
   rc <- (1-pweibull(t,v_alpha, v_sigma)) *
      (1-pweibull(t,s_alpha, s_sigma))
   return(rc)
  }


#mean survival time for serologic and virologic. 
mns$sero.viro <- NA
for (i in 1:length(chains$lm)) {
  mns$sero.viro[i] <- 
    integrate(to.integrate, 0,Inf, 
              v_alpha= chains$v_alpha[i],
              v_sigma= chains$v_sigma[i],
              s_alpha= chains$s_alpha[i],
              s_sigma= chains$s_sigma[i])$value
  }

symptom.relative.sero <- quantile(mns$sero.viro/mns$viro,
                                  prob=c(0.025, .5, 0.975))


symptom.relative.inc
symptom.relative.sero


```


#Single interval censored survreg analysis.

Does a single interval censored analysis using survreg

```{r}
                   
#do incunation period first.
inc.low <- pmax(zika.anal$SL - zika.anal$ER, 0.01)
inc.high <-zika.anal$SR - zika.anal$EL
inc.surv <- Surv(inc.low, inc.high, event=rep(3, length(inc.high)),type="interval")

inc.sr <- survreg(inc.surv~1, dist="lognormal")
inc.sr
confint(inc.sr)


#do viral shedding
v.low <-  pmax(zika.anal$VCL - zika.anal$ER, 0.01)
v.high <-zika.anal$VCR - zika.anal$EL
v.surv <- Surv(v.low, v.high, type="interval2")
v.sr <- survreg(v.surv~1, dist="weibull", scale=1/vshed.fit.stan.LWW[1,1])
v.sr
confint(v.sr)

#do serology
s.low <-  pmax(zika.anal$ACL - zika.anal$ER, 0.01)
s.high <-zika.anal$ACR - zika.anal$EL
s.surv <- Surv(s.low, s.high, type="interval2")
s.sr <- survreg(s.surv~1, dist="weibull", scale =1/sero.fit.stan.LWW[1,1])
s.sr
confint(s.sr)

```


#Combined stan analysis Lognomal for incubation period, Gamma for others

```{r, eval=FALSE}
require(rstan)


#Create a dataset with everythign lined up for passing into STAN
zika.anal <- zika.inc[,c("UID","EL","ER","SL","SR")]
zika.anal <- merge(zika.anal, vshed.bl.anal, all=TRUE)
zika.anal$type <- 0 #we don't use this, but helps with the merge.
zika.anal <- merge(zika.anal, sero.anal, all=TRUE)

zika.anal$SL[which(is.na(zika.anal$SL))] <- 
  zika.anal$EL[which(is.na(zika.anal$SL))]
zika.anal$SR[which(is.na(zika.anal$SR))] <- 
   zika.anal$ER[which(is.na(zika.anal$SR))] + 100

zika.anal$ACL[which(is.na(zika.anal$ACL))] <-
  zika.anal$EL[which(is.na(zika.anal$ACL))]
zika.anal$ACR[which(is.na(zika.anal$ACR))] <- 
  zika.anal$ER[which(is.na(zika.anal$ACR))] + 720

zika.anal$VCL[which(is.na(zika.anal$VCL))] <-
  zika.anal$EL[which(is.na(zika.anal$VCL))]
zika.anal$VCR[which(is.na(zika.anal$VCR))] <- 
  zika.anal$ER[which(is.na(zika.anal$VCR))] + 720


zika.anal$VCR[which(is.infinite(zika.anal$VCR))] <- 
  zika.anal$ER[which(is.infinite(zika.anal$VCR))] + 720

#renormalize everything so EL is 0
zika.anal[,c("EL","ER","SL","SR","VCL","VCR","ACL","ACR")] <-
  zika.anal[,c("EL","ER","SL","SR","VCL","VCR","ACL","ACR")] -zika.anal$EL  


#for STAN purposes, it is usefule if all of these variables index 
#from 0
stan.data <- list(N=nrow(zika.anal),
                  EL=zika.anal$EL,
                  ER=zika.anal$ER,
                  SL=zika.anal$SL,
                  SR=zika.anal$SR,
                  VL=zika.anal$VCL,
                  VR=zika.anal$VCR, 
                  AL=zika.anal$ACL,
                  AR=zika.anal$ACR)


#initialization version for the lognormal/weibull stan model


#use uniform from halfway trhough the exposure interval to ER.
#still fails occasionally, but MUCH less often. 
stan.init.LGG <- function() {
    rc <-list(E=runif(stan.data$N, (stan.data$EL+stan.data$ER)/2, stan.data$ER),
                lm= runif(1,log(2),log(10)),
                lsd = runif(1,.1,2),
                v_alpha = 5,#,runif(1,5,10),
                v_beta= .1,# runif(1,.5,2) 
                s_alpha = 5,#,runif(1,5,10),
                s_beta= .1# runif(1,.5,2) 
                )
    #print(rc) #DEBUG
    #curve(dweibull(x,rc$v_alpha, rc$v_sigma), from=0, to=50) #DEBUG
    return(rc)
    }



#Need at least 35,000 iterations to converge
full.fit.LGG <- stan("DistributionFitLGG.stan", data=stan.data,
                  init=stan.init.LGG,
                  iter=100, chains=4, thin=1)

save(full.fit.LGG, file = "../FullFit.LGG.RData")

```


#Combined stan analysis Lognomal for incubation period, Gamma others
#THIS DOES NOT WORK!!!!
This is not working at this point.
```{r, eval=FALSE}
require(rstan)


#Create a dataset with everythign lined up for passing into STAN
zika.anal <- zika.inc[,c("UID","EL","ER","SL","SR")]
zika.anal <- merge(zika.anal, vshed.bl.anal, all=TRUE)

zika.anal$SL[which(is.na(zika.anal$SL))] <- 
  zika.anal$EL[which(is.na(zika.anal$SL))]
zika.anal$SR[which(is.na(zika.anal$SR))] <- 
   zika.anal$ER[which(is.na(zika.anal$SR))] + 100

zika.anal$VCL[which(is.na(zika.anal$VCL))] <-
  zika.anal$EL[which(is.na(zika.anal$VCL))]
zika.anal$VCR[which(is.na(zika.anal$VCR))] <- 
  zika.anal$ER[which(is.na(zika.anal$VCR))] + 720

zika.anal$VCR[which(is.infinite(zika.anal$VCR))] <- 
  zika.anal$ER[which(is.infinite(zika.anal$VCR))] + 720

#renormalize everything so EL is 0
zika.anal[,c("EL","ER","SL","SR","VCL","VCR")] <-
  zika.anal[,c("EL","ER","SL","SR","VCL","VCR")] -zika.anal$EL  


#for STAN purposes, it is usefule if all of these variables index 
#from 0
stan.data <- list(N=nrow(zika.anal),
                  EL=zika.anal$EL,
                  ER=zika.anal$ER,
                  SL=zika.anal$SL,
                  SR=zika.anal$SR,
                  VL=zika.anal$VCL,
                  VR=zika.anal$VCR)


#initialization version for the lognormal/gamma stan model
#in stan alpha = shape in R and beta = rate in R

#use uniform from halfway trhough the exposure interval to ER.
#still fails occasionally, but MUCH less often. 
stan.init.LG <- function() {
    rc <-list(E=runif(stan.data$N, (stan.data$EL+stan.data$ER)/2, stan.data$ER),
                lm= runif(1,log(2),log(10)),
                lsd = runif(1,.1,2),
                v_alpha = 5,#runif(1,5,10),
                v_beta= .1#runif(1,.1,1) 
                )
    #print(rc) #DEBUG
    
    return(rc)
    }



#Need at least 20,000 iterations to converge
full.fit <- stan("DistributionFitLG.stan", data=stan.data,
                  init=stan.init.LG,
                  iter=100, chains=4)


#now we need to calculate the quantiles from the chains and construct our
#data frame
chains <- extract(full.fit)


inc.fit.stan <- c(mean(chains$lm), quantile(chains$lm,prob=c(0.025,0.975)))
inc.fit.stan <- rbind(inc.fit.stan,
        c(mean(chains$lsd), quantile(chains$lsd,prob=c(0.025,0.975))))
for (q in c(0.05, 0.25, 0.5, 0.75, 0.95)) {
  tmp <- qlnorm(q, chains$lm, chains$lsd)
  inc.fit.stan <- rbind(inc.fit.stan,
                        c(mean(tmp), quantile(tmp, prob=c(0.025, 0.975))))
}

colnames(inc.fit.stan) <- c("est","CIlow","CIhigh")
rownames(inc.fit.stan) <- c("meanlog",
                            "sdlog",
                            "p5","p25","p50","p75","p95")



#in stan alpha = shape in R and beta = rate in R
vshed.fit.stan <- c(mean(chains$v_alpha), 
                    quantile(chains$v_alpha,prob=c(0.025,0.975)))
vshed.fit.stan <- rbind(vshed.fit.stan,
        c(mean(chains$v_beta), 
          quantile(chains$v_beta,prob=c(0.025,0.975))))
for (q in c(0.05, 0.25, 0.5, 0.75, 0.95)) {
  tmp <- qgamma(q, shape=chains$v_alpha, rate=chains$v_beta)
  vshed.fit.stan <- rbind(vshed.fit.stan,
                        c(mean(tmp), quantile(tmp, prob=c(0.025, 0.975))))
}

colnames(vshed.fit.stan) <- c("est","CIlow","CIhigh")
rownames(vshed.fit.stan) <- c("alpha (shape)",
                            "beta (rate)",
                            "p5","p25","p50","p75","p95")


```




#Appendix

##Fitting just incubation period and time to seroconversion with Lognormal and gamma

```{r, eval=FALSE}
require(rstan)


#Create a dataset with everythign lined up for passing into STAN
zika.anal <- zika.inc[,c("UID","EL","ER","SL","SR")]
zika.anal <- merge(zika.anal, vshed.bl.anal, all=TRUE)

zika.anal$SL[which(is.na(zika.anal$SL))] <- 
  zika.anal$EL[which(is.na(zika.anal$SL))]
zika.anal$SR[which(is.na(zika.anal$SR))] <- 
   zika.anal$ER[which(is.na(zika.anal$SR))] + 100

zika.anal$VCL[which(is.na(zika.anal$VCL))] <-
  zika.anal$EL[which(is.na(zika.anal$VCL))]
zika.anal$VCR[which(is.na(zika.anal$VCR))] <- 
  zika.anal$ER[which(is.na(zika.anal$VCR))] + 720

zika.anal$VCR[which(is.infinite(zika.anal$VCR))] <- 
  zika.anal$ER[which(is.infinite(zika.anal$VCR))] + 720

#renormalize everything so EL is 0
zika.anal[,c("EL","ER","SL","SR","VCL","VCR")] <-
  zika.anal[,c("EL","ER","SL","SR","VCL","VCR")] -zika.anal$EL  


#for STAN purposes, it is usefule if all of these variables index 
#from 0
stan.data <- list(N=nrow(zika.anal),
                  EL=zika.anal$EL,
                  ER=zika.anal$ER,
                  SL=zika.anal$SL,
                  SR=zika.anal$SR,
                  VL=zika.anal$VCL,
                  VR=zika.anal$VCR)


#initialization version for the lognormal/weibull stan model


#use uniform from halfway trhough the exposure interval to ER.
#still fails occasionally, but MUCH less often. 
stan.init.LW <- function() {
    rc <-list(E=runif(stan.data$N, (stan.data$EL+stan.data$ER)/2, stan.data$ER),
                lm= runif(1,log(2),log(10)),
                lsd = runif(1,.1,2),
                v_alpha = 2,#,runif(1,5,10),
                v_sigma= 50# runif(1,.5,2) 
                )
    print(rc) #DEBUG
    curve(dweibull(x,rc$v_alpha, rc$v_sigma), from=0, to=50)
    return(rc)
    }



#Need at least 20,000 iterations to converge
full.fit.LW <- stan("DistributionFitLW.stan", data=stan.data,
                  init=stan.init.LW,
                  iter=35000, chains=4, thin=10)


#now we need to calculate the quantiles from the chains and construct our
#data frame
chains <- extract(full.fit)


inc.fit.stan.LW <- c(mean(chains$lm), quantile(chains$lm,prob=c(0.025,0.975)))
inc.fit.stan.LW <- rbind(inc.fit.stan.LW,
        c(mean(chains$lsd), quantile(chains$lsd,prob=c(0.025,0.975))))
for (q in c(0.05, 0.25, 0.5, 0.75, 0.95)) {
  tmp <- qlnorm(q, chains$lm, chains$lsd)
  inc.fit.stan.LW <- rbind(inc.fit.stan.LW,
                        c(mean(tmp), quantile(tmp, prob=c(0.025, 0.975))))
}

colnames(inc.fit.stan.LW) <- c("est","CIlow","CIhigh")
rownames(inc.fit.stan.LW) <- c("meanlog",
                            "sdlog",
                            "p5","p25","p50","p75","p95")



#in stan alpha = shape in R and beta = rate in R
vshed.fit.stan.LW <- c(mean(chains$v_alpha), 
                    quantile(chains$v_alpha,prob=c(0.025,0.975)))
vshed.fit.stan.LW <- rbind(vshed.fit.stan.LW,
        c(mean(chains$v_sigma), 
          quantile(chains$v_sigma,prob=c(0.025,0.975))))
for (q in c(0.05, 0.25, 0.5, 0.75, 0.95, 0.99)) {
  tmp <- qweibull(q, chains$v_alpha, chains$v_sigma)
  vshed.fit.stan.LW <- rbind(vshed.fit.stan.LW,
                        c(mean(tmp), quantile(tmp, prob=c(0.025, 0.975))))
}

colnames(vshed.fit.stan.LW) <- c("est","CIlow","CIhigh")
rownames(vshed.fit.stan.LW) <- c("alpha (shape)",
                            "sigma (scale)",
                            "p5","p25","p50","p75","p95","p99")


```


## Some leftover plot code
```{r}

##########CODE FOR MAKING AN UGLY COMBINED PLOT########
#Try to use ggplot to plot a beautiful display of the data. 

#first make data frames to hold everything
inc.curve <- NULL
vc.curve <- NULL
sero.curve<- NULL

for (q in seq(0,30,.1)) {
    tmp <- plnorm(q, chains$lm, chains$lsd)
    tmp <- quantile(tmp, prob=c(0.025, .5, 0.975))
    inc.curve <- rbind(inc.curve, c(q=q,
                                    plow=tmp[1],
                                    pmid=tmp[2],
                                    phigh=tmp[3]))
    
    tmp <- 1-pweibull(q, chains$v_alpha, chains$v_sigma)
    tmp <- quantile(tmp, prob=c(0.025, .5, 0.975))
    vc.curve <- rbind(vc.curve, c(q=q,
                                  plow=tmp[1],
                                  pmid=tmp[2],
                                  phigh=tmp[3]))
    
    tmp <- pweibull(q, chains$s_alpha, chains$s_sigma)
    tmp <- quantile(tmp, prob=c(0.025, .5, 0.975))
    sero.curve <- rbind(sero.curve, c(q=q,
                                  plow=tmp[1],
                                  pmid=tmp[2],
                                  phigh=tmp[3]))
}
inc.curve <- as.data.frame(inc.curve)
colnames(inc.curve) <- c("q","incplow","incpmid","incphigh")

vc.curve <- as.data.frame(vc.curve)
colnames(vc.curve) <- c("q","vcplow","vcpmid","vcphigh")

sero.curve <- as.data.frame(sero.curve)
colnames(sero.curve) <- c("q","scplow","scpmid","scphigh")


require(ggplot2)


comb.curve<-merge(merge(inc.curve, vc.curve), sero.curve)
comb.plt <- ggplot(comb.curve, aes(x=q)) +
  geom_ribbon(aes(ymin=incplow, ymax=incphigh), fill="blue", alpha=.2) +
  geom_line(aes(y=incpmid), col="blue") +
  geom_ribbon(aes(ymin=vcplow, ymax=vcphigh), fill="red", alpha=.2) +
  geom_line(aes(y=vcpmid), col="red") +
  geom_ribbon(aes(ymin=scplow, ymax=scphigh), fill="green", alpha=.2) +
  geom_line(aes(y=scpmid), col="green") + xlim(0,25)

comb.plt

```


##Combined stan analysis Lognormal for all distributions
```{r}
library(rstan)


#Create a dataset with everythign lined up for passing into STAN
zika.anal <- zika.inc[,c("UID","EL","ER","SL","SR")]
zika.anal <- merge(zika.anal, vshed.bl.anal, all=TRUE)

zika.anal$SL[which(is.na(zika.anal$SL))] <- 
  zika.anal$EL[which(is.na(zika.anal$SL))]
zika.anal$SR[which(is.na(zika.anal$SR))] <- 
   zika.anal$ER[which(is.na(zika.anal$SR))] + 100

zika.anal$VCL[which(is.na(zika.anal$VCL))] <-
  zika.anal$EL[which(is.na(zika.anal$VCL))]
zika.anal$VCR[which(is.na(zika.anal$VCR))] <- 
  zika.anal$ER[which(is.na(zika.anal$VCR))] + 720

zika.anal$VCR[which(is.infinite(zika.anal$VCR))] <- 
  zika.anal$ER[which(is.infinite(zika.anal$VCR))] + 720

#Replace

#for STAN purposes, it is usefule if all of these variables index 
#from 0
stan.data <- list(N=nrow(zika.anal),
                  EL=zika.anal$EL,
                  ER=zika.anal$ER,
                  SL=zika.anal$SL,
                  SR=zika.anal$SR,
                  VL=zika.anal$VCL,
                  VR=zika.anal$VCR)


#initialization version for the lognormal/lognormal stan model
stan.init.LL <- function() {
    rc <-list(E=runif(stan.data$N, stan.data$EL, stan.data$ER),
                lm=runif(1,log(2),log(10)),
                lsd = runif(1,.1,2),
                v_lm = runif(1,log(2),log(10)),
                v_lsd = runif(1,.1,2) 
                )
    print(rc)
    return(rc)
    }



#Need at least 20,000 iterations to converge
full.fit.LL <- stan("DistributionFitLL.stan", data=stan.data,
                  init=stan.init.LL,
                  iter=35000, chains=4, thin=10)


#now we need to calculate the quantiles from the chains and construct our
#data frame
chains <- extract(full.fit.LL)


inc.fit.stanLL <- c(mean(chains$lm), quantile(chains$lm,prob=c(0.025,0.975)))
inc.fit.stanLL <- rbind(inc.fit.stanLL,
        c(mean(chains$lsd), quantile(chains$lsd,prob=c(0.025,0.975))))
for (q in c(0.05, 0.25, 0.5, 0.75, 0.95)) {
  tmp <- qlnorm(q, chains$lm, chains$lsd)
  inc.fit.stanLL <- rbind(inc.fit.stanLL,
                        c(mean(tmp), quantile(tmp, prob=c(0.025, 0.975))))
}

colnames(inc.fit.stanLL) <- c("est","CIlow","CIhigh")
rownames(inc.fit.stanLL) <- c("meanlog",
                            "sdlog",
                            "p5","p25","p50","p75","p95")



vshed.fit.stanLL <- c(mean(chains$v_lm), quantile(chains$v_lm,prob=c(0.025,0.975)))
vshed.fit.stanLL <- rbind(vshed.fit.stanLL,
        c(mean(chains$v_lsd), quantile(chains$v_lsd,prob=c(0.025,0.975))))
for (q in c(0.05, 0.25, 0.5, 0.75, 0.95)) {
  tmp <- qlnorm(q, chains$v_lm, chains$v_lsd)
  vshed.fit.stanLL <- rbind(vshed.fit.stanLL,
                        c(mean(tmp), quantile(tmp, prob=c(0.025, 0.975))))
}

colnames(vshed.fit.stanLL) <- c("est","CIlow","CIhigh")
rownames(vshed.fit.stanLL) <- c("meanlog",
                            "sdlog",
                            "p5","p25","p50","p75","p95")


```


